### 2025-01-21 documents/paper.tex
Presentation slides.

\documentclass[11pt, a4paper, leqno]{article}
\usepackage{a4wide}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{float, afterpage, rotating, graphicx}
\usepackage{epstopdf}
\usepackage{longtable, booktabs, tabularx}
\usepackage{fancyvrb, moreverb, relsize}
\usepackage{eurosym, calc}
% \usepackage{chngcntr}
\usepackage{amsmath, amssymb, amsfonts, amsthm, bm}
\usepackage{caption}
\usepackage{mdwlist}
\usepackage{xfrac}
\usepackage{setspace}
\usepackage[dvipsnames]{xcolor}
\usepackage{subcaption}
\usepackage{minibox}
% \usepackage{pdf14} % Enable for Manuscriptcentral -- can't handle pdf 1.5
% \usepackage{endfloat} % Enable to move tables / figures to the end. Useful for some
% submissions.

\usepackage[unicode=true]{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    anchorcolor=black,
    citecolor=NavyBlue,
    filecolor=black,
    menucolor=black,
    runcolor=black,
    urlcolor=NavyBlue
}

\widowpenalty=10000
\clubpenalty=10000

\setlength{\parskip}{1ex}
\setlength{\parindent}{0ex}
\setstretch{1.5}

\begin{document}

\title{Reproduction and Expansion of \\"SPECIALIZATION TRENDS IN ECONOMICS RESEARCH" by Sebastian Galiani, Ramiro H. Gálvez, and Ian Nachman}

\author{Felix Schmitz}

\date{
    {\bf Under construction}
    \\[1ex]
    \today
}

\maketitle
\footnotetext[1]{Email: \href{mailto:s87fschm@uni-bonn.de}{s87fschm@uni-bonn.de}}

\begin{abstract}
    Some abstract here.
\end{abstract}

\clearpage

\end{document}

### 2025-01-21 documents/presentation.tex
Presentation slides.

\documentclass[aspectratio=1610]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{bbm}

\usepackage[
    natbib=true,
    bibencoding=inputenc,
    bibstyle=authoryear-ibid,
    citestyle=authoryear-comp,
    maxcitenames=3,
    maxbibnames=10,
    useprefix=false,
    sortcites=true,
    backend=biber
]{biblatex}

\title{Specialization Trends in Economics Research\\
using JEL Codes}
\subtitle{Research Proposal Presentation for Topics Course Text Data in Economics}
\date{January 21, 2025}
\author{Felix Schmitz}

\usetheme{UniBonn}

\begin{document}
\usebeamertemplate{custom equation spacing}

\begin{frame}[plain]
	\titlepage
\end{frame}

\begin{framecontent}
	\frametitle{Table of contents}
\end{framecontent}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
	\frametitle{Motivation}
	\begin{center}
		\bfseries{
			Are subfields in economics research becoming more specialized?\\
			Are there new subfields emerging?\\
			Are Journal of Economic Literature (JEL) codes of publications helpful in answering these questions?
		}
	\end{center}

	To answer these questions, I will:
	\begin{itemize}
		\item Analyze the use of individual JEL codes in economics research over time
		\item Inspect the combinations of JEL codes used (combinations vanishing/emerging)
		\item Compare results to other research methods analyzing specialization trends
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Motivation}
	\framesubtitle{A primer on JEL codes}
	\begin{itemize}
		\item Journal of Economic Literature: quarterly published by the American Economic Association
		\item JEL Codes are a classification system for research articles widely used in Economics
		\item Hierarchical structure with single character and two-digit codes
		\item Self-classification by authors (may be adapted during review process)
		\item Categories range from general economics to specific subfields, but do not cover methods
		\item Example follows
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Motivation}
	\framesubtitle{Example: Galiani (2023) classified as JEL A11, A14}
	\textbf{A. General Economics and Teaching}
    \begin{itemize}
        \item A1 General Economics
        \begin{itemize}
            \item A10 General
            \item \textbf{A11 Role of Economics • Role of Economists • Market for Economists}
            \item A12 Relation of Economics to Other Disciplines
            \item A13 Relation of Economics to Social Values
            \item \textbf{A14 Sociology of Economics}
            \item A19 Other
        \end{itemize}
        \item A2 Economic Education and Teaching of Economics
        \begin{itemize}
            \item A20 General
            \item A21 Pre-college
            \item $\cdots$
        \end{itemize}
        \item A3 Collective Works
        \begin{itemize}
            \item A30 General
            \item $\cdots$
        \end{itemize}
    \end{itemize}
	\textbf{B. History of Economic Thought, Methodology, and Heterodox Approaches}	
\end{frame}

\begin{frame}
	\frametitle{Motivation}
	\framesubtitle{Literature}
	\begin{itemize}
		\item Heikkila (2022): History of JEL codes and potential research applications
		\item Heikkila (2024): Uses abstracts and LLMs to create (artificial) JEL codes for comparison
		\item Kelly (2011): Analyze the publication of subjects, the development of JEL codes, and define specialty journals using JEL codes (1969-2007)
		\item Rath (2016): Build upon the work of Kelly (2011) and analyze general JEL code use developments (add 2007-2013)
		\item Angrist (2017) and Angrist (2020): NLP techniques to assign specific field of economics research tags
		\item Galiani (2023) (first presentation): Used both hand-coded and machine-coded articles to analyze specialization trends in fields of economics research
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Sources}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
	\frametitle{Data Sources}
	\framesubtitle{already acquired}

	\textbf{Constellate}
	\begin{itemize}
		\item Metadata (title, DOI, author names, publication date, journal of publication)
		\item unigrams, bigrams, and trigrams
	\end{itemize}

	\textbf{Semantic Scholar}
	\begin{itemize}
		\item Metadata (title, publication date, identifiers, authors, fields of study tags)
		\item Abstracts
		\item Comprehensive citation and reference data
	\end{itemize}

	\textbf{Galiani (2016) and Galiani (2020)}
	\begin{itemize}
		\item Manually tagged field of economics research
		\item Include: applied, applied theory, econometric methods, theory
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Data Sources}
	\framesubtitle{initial sample}

	\textbf{Research Papers in Economics (RePEc) IDEAS}
	\begin{itemize}
		\item Metadata (title, author, publication date, journal of publication)
		\item JEL codes
		\item Accessible via FTP server (connecting to all publisher FTP servers providing the metadata)
		\item A Python package for accessing RePEc data is available on \href{https://github.com/andrei-dubovik/repec}{GitHub} (not used yet)
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Data Sources}
		\framesubtitle{challenges so far}

		\begin{itemize}
			\item \textbf{Data availability:} Open access limited, especially full dataset download
			\begin{itemize}
				\item Constellate will be discontinued in June
				\item Semantic Scholar broader coverage but less detailed metadata
				\item RePEc discourages scraping, no public API, no documentation of dataset building
				\item EconLit (American Economics Association) behind paywall
			\end{itemize}
			\item \textbf{Size of dataset:} 
			\begin{itemize}
				\item APIs are rare, hence datasets contain large number of variables
				\item Constellate and Semantic Scholar sample from Galiani (2023) are 6GB (each) in zipped format
			\end{itemize}
			\item \textbf{Data quality:}
			\begin{itemize}
				\item JEL codes are self-assigned by authors
				\item JEL codes are not always assigned
				\item JEL codes may reflect the content poorly
			\end{itemize}
		\end{itemize}
	\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
	\frametitle{Methodology}
	\begin{itemize}
		\item \textbf{Trends of individual JEL codes used:}
		\begin{itemize}
			\item Publication counts of individual JEL codes over time
			\item Average number of JEL codes per article
			\item Analysis of the above in combination with journals
		\end{itemize}

		\item \textbf{Cluster analysis of JEL code combinations:}
		\begin{itemize}
			\item Clustering of JEL code combinations (e.g. via Louvain method for community detection)
			\item Analysis of clusters over time
		\end{itemize}

		\item \textbf{Comparison to other research methods:}
		\begin{itemize}
			\item Use of text data methods will be highly relevant here
			\item Topic modeling via n-grams, LDA, or available data sources
			\item Comparison to other research fields via Constellate or Semantic Scholar data
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Methodology}
	\framesubtitle{A primer on Louvain method}
	\begin{itemize}
		\item Goal is to find non-overlapping communities inside a large network
		\item Nodes are individual JEL codes, and edges represent the strength of their co-occurrence
		\item Iterative process of optimizing modularity $Q$ via a weighted graph
		\begin{equation}
			Q = \frac{1}{2m} \sum_{i=1}^N \sum_{j=1}^N \left( A_{ij} - \frac{k_i k_j}{2m} \right) \delta(c_i, c_j)
		\end{equation}
		\item $A_{ij}$: edge weight between nodes $i$ and $j$
		\item $k_i$ and $k_j$: sum of weights of edges connected to each nodes $i$ and $j$ 
		\item $m$: sum of all edge weights
		\item $N$ number of nodes
		\item $c_i$ and $c_j$: community assignment of nodes $i$ and $j$
		\item $\delta(c_i, c_j) = \mathbbm{1}[c_i = c_j]$: Kronecker delta function
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Methodology}
	\framesubtitle{A primer on Louvain method (continued)}
	\begin{itemize}
		\item Modularity $Q$ of community $c$ then becomes:
		\begin{equation}
			Q_c = \frac{1}{2m} \sum_{i=1}^N \sum_{j=1}^N  A_{ij} \mathbbm{1}{c_i = c_j = c} - \left(\sum_{i=1}^N \frac{k_i }{2m} \mathbbm{1}{c_i = c} \right)^2
			= \frac{\Sigma_{in}}{2m} - \left(\frac{\Sigma_{tot}}{2m}\right)^2
		\end{equation}
		\item $\Sigma_{in}$: sum of the weights of the edges inside the community
		\item $\Sigma_{tot}$: sum of the weights of the edges connected to nodes in the community (including to nodes outside the community)
		\item Since the method provides non-overlapping communities:
		\begin{equation}
			Q = \sum_{c=1}^N Q_c
		\end{equation}
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary \& Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
	\frametitle{Summary \& Discussion}
	\begin{itemize}
		\item Research so far focused on text analysis and not on JEL code combinations
		\item Access to JEL codes so far limited, but feasible
		\item Weakness of JEL codes in representing content and correct assignment
		\item Methodology: Louvain method for community detection (use the information of single JEL code for clustering combinations?)
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Summary \& Discussion}
	\framesubtitle{What else?}

	\begin{center}
		\textbf{Any ideas for understanding specialization trends?\\
		Do you have any suggestions for the methodology?\\
		Any other comments or ideas?}
	\end{center}
\end{frame}

\begin{frame}
    \frametitle{A graphical representation of the Louvain method (Wikipedia)}
    \begin{columns}
        \begin{column}{0.33\textwidth}
            \includegraphics[width=\textwidth]{Louvain_Step1.png}
			\captionof{figure}{Step 1}
			Node initialization with singleton community
        \end{column}

        \begin{column}{0.33\textwidth}
            \includegraphics[width=\textwidth]{Louvain_Step2.png}
			\captionof{figure}{Step 2}
			Community assignment based on modularity gain
        \end{column}

        \begin{column}{0.33\textwidth}
            \includegraphics[width=\textwidth]{Louvain_Step3.png}
			\captionof{figure}{Step 3}
			Community aggregation and edge weight update
        \end{column}
    \end{columns}
\end{frame}
\end{document}

### 2025-01-21 documents/task.py

@@ -1,0 +1,33 @@
+"""Tasks for compiling the paper and presentation(s)."""
+
+import shutil
+
+import pytask
+from pytask_latex import compilation_steps as cs
+
+from econ_spec_jel.config import BLD, DOCUMENTS, ROOT
+
+documents = ["paper", "presentation"]
+
+for document in documents:
+
+    @pytask.mark.latex(
+        script=DOCUMENTS / f"{document}.tex",
+        document=BLD / "documents" / f"{document}.pdf",
+        compilation_steps=cs.latexmk(
+            options=("--pdf", "--interaction=nonstopmode", "--synctex=1", "--cd"),
+        ),
+    )
+    @pytask.task(id=document)
+    def task_compile_document():
+        """Compile the document specified in the latex decorator."""
+
+    kwargs = {
+        "depends_on": BLD / "documents" / f"{document}.pdf",
+        "produces": ROOT / f"{document}.pdf",
+    }
+
+    @pytask.task(id=document, kwargs=kwargs)
+    def task_copy_to_root(depends_on, produces):
+        """Copy a document to the root directory for easier retrieval."""
+        shutil.copy(depends_on, produces)

### 2025-01-21 src/econ_spec_jel/data_scraping/task.py
Initial commit of scraping.

@@ -1,0 +1,15 @@
+""" Task functions for data scraping. """
+# write a loop over the two tasks and pass the number to each of them?
+# how else could I achieve parallelization/be sure not to need to re-run them all?
+
+def task_scrape_metadata():
+    # base url: https://www.iza.org/publications/dp/[number]
+    # numbers range from 1 to 17,628 (as of 21.01.2025)
+    # there only 17,622 papers though
+    # for each scrape, author(s), month, year, title, abstract, keywords, jel codes
+    return None
+
+def task_scrape_paper():
+    # base url: https://docs.iza.org/dp[number].pdf
+    # download the pdf file
+    return None

### 2025-02-13 src/econ_spec_jel/config.py
Metadata and file scraping.

@@ -1,5 +1,6 @@
 """Contains the general configuration of the project."""
 
+from typing import NamedTuple
 from pathlib import Path
 
 from pytask import DataCatalog
@@ -10,9 +11,25 @@ BLD = ROOT.joinpath("bld").resolve()
 DATA = ROOT.joinpath("data").resolve()
 DOCUMENTS = ROOT.joinpath("documents").resolve()
 
-DATASOURCES = ["constellate", "semantic_scholar"]
+DATACATALOGS = {"metadata": DataCatalog(name="metadata"),
+                "files": DataCatalog(name="files")}
 
-DATACATALOG = {datasource: DataCatalog(name=datasource) for datasource in DATASOURCES}
+MAX_DP_NUMBER = 17680
 
 
-__all__ = ["BLD", "DATA", "DATACATALOG", "DATASOURCES", "DOCUMENTS", "ROOT", "SRC"]
+class Scraper(NamedTuple):
+    """ NamedTuple to store the base urls and the dp number of a discussion paper."""
+    metadata_base_url: str
+    file_base_url: str
+    dp_number: int
+
+SCRAPERS = [
+    Scraper(
+        metadata_base_url="https://www.iza.org/publications/dp/",
+        file_base_url="https://docs.iza.org/dp",
+        dp_number=num
+        )
+    for num in range(1, MAX_DP_NUMBER+1)
+    ]
+
+__all__ = ["BLD", "DATA", "DATACATALOGS", "DOCUMENTS", "MAX_DP_NUMBER", "ROOT", "Scraper", "SCRAPERS", "SRC"]

### 2025-02-13 src/econ_spec_jel/data_scraping/helper.py

@@ -1,0 +1,50 @@
+from bs4 import BeautifulSoup
+import requests
+
+def extract_metadata(response: requests.Response, dp_number: int) -> dict:
+    """ Extract metadata from the response of a discussion paper page.
+
+    Args:
+        response (requests.Response): Response object of the discussion paper page.
+        dp_number (int): Discussion paper number.
+
+    Returns:
+        dict: Metadata.
+    """
+    soup = BeautifulSoup(response.text, "html.parser")
+
+    title = soup.select_one('meta[property="og:title"]')['content'].strip()
+    pub_date = soup.select_one('div.publications-header p').text.strip()
+    pub_month, pub_year = pub_date.split(" ")
+    authors_information = soup.select('div.authors a')
+    author_names = [a.text.strip() for a in authors_information]
+    author_urls = [a['href'] for a in authors_information]
+    abstract = soup.select_one('div.element-copyexpandable p').text.strip()
+    keywords = [
+        li.text.strip()
+        for h3 in soup.find_all('h3', class_='box-list-headline')
+        if 'Keywords' in h3.text
+        for li in h3.find_next('ul').find_all('li')
+    ]
+    jel_codes = [
+        li.text.strip()
+        for h3 in soup.find_all('h3', class_='box-list-headline')
+        if 'JEL Codes' in h3.text
+        for li in h3.find_next('ul').find_all('li')
+    ]
+    try:
+        file_url = soup.select_one('a.download-link')['href']
+    except TypeError:
+        file_url = None
+    return {
+        'dp_number': dp_number,
+        'title': title,
+        'author_names': author_names,
+        'author_urls': author_urls,
+        'publication_date_month': pub_month,
+        'publication_date_year': pub_year,
+        'abstract': abstract,
+        'keywords': keywords,
+        'jel_codes': jel_codes,
+        'file_url': file_url
+    }

### 2025-02-13 src/econ_spec_jel/data_scraping/task.py

@@ -1,15 +1,56 @@
 """ Task functions for data scraping. """
 # write a loop over the two tasks and pass the number to each of them?
 # how else could I achieve parallelization/be sure not to need to re-run them all?
+import pytask
+from pytask import task
+import requests
+from typing import Annotated
+from pathlib import Path
 
-def task_scrape_metadata():
-    # base url: https://www.iza.org/publications/dp/[number]
-    # numbers range from 1 to 17,628 (as of 21.01.2025)
-    # there only 17,622 papers though
-    # for each scrape, author(s), month, year, title, abstract, keywords, jel codes
-    return None
+from econ_spec_jel.config import DATACATALOGS, Scraper, SCRAPERS
+from econ_spec_jel.data_scraping.helper import extract_metadata
 
-def task_scrape_paper():
-    # base url: https://docs.iza.org/dp[number].pdf
-    # download the pdf file
-    return None
+
+def _scrape_metadata(base_url: str, dp_number: int) -> dict:
+    url = f"{base_url}{dp_number}"
+    response = requests.get(url)
+    if response.status_code != 200:
+        return {'dp_number': dp_number} | {k: None for k in ['title', 'author_names', 'author_urls', 'publication_date_month', 'publication_date_year', 'abstract', 'keywords', 'jel_codes', 'file_url']}
+    return extract_metadata(response, dp_number)
+
+def _download_file(base_url: str, dp_number: int) -> bytes:
+    url = f"{base_url}{dp_number}.pdf"
+    response = requests.get(url)
+    return response.content
+
+
+for scraper in SCRAPERS:
+    if not DATACATALOGS["metadata"][str(scraper.dp_number)].path.is_file():
+        @task(id=str(scraper.dp_number))
+        def task_scrape_metadata(
+            scraper: Scraper=scraper,
+            ) -> Annotated[Path, DATACATALOGS["metadata"][str(scraper.dp_number)]]:
+            """ Scrape metadata from discussion paper page.
+
+            Args:
+                scraper: Scraper object with metadata_base_url and dp_number attributes.
+
+            Returns:
+                dict: Metadata.
+            """
+            return _scrape_metadata(base_url=scraper.metadata_base_url, dp_number=scraper.dp_number)
+
+    if (not DATACATALOGS["files"][str(scraper.dp_number)].path.is_file()) and (DATACATALOGS["metadata"][str(scraper.dp_number)].load()["file_url"] is not None):
+        @task(id=str(scraper.dp_number))
[…]

### 2025-02-13 documents/task.py
pre-commit hooks.

@@ -1,11 +1,14 @@
 """Tasks for compiling the paper and presentation(s)."""
 
 import shutil
+from pathlib import Path
 
 import pytask
 from pytask_latex import compilation_steps as cs
 
-from econ_spec_jel.config import BLD, DOCUMENTS, ROOT
+from econ_spec_jel.config import BLD
+from econ_spec_jel.config import DOCUMENTS
+from econ_spec_jel.config import ROOT
 
 documents = ["paper", "presentation"]
 
@@ -19,7 +22,7 @@ for document in documents:
         ),
     )
     @pytask.task(id=document)
-    def task_compile_document():
+    def task_compile_document() -> None:
         """Compile the document specified in the latex decorator."""
 
     kwargs = {
@@ -28,6 +31,6 @@ for document in documents:
     }
 
     @pytask.task(id=document, kwargs=kwargs)
-    def task_copy_to_root(depends_on, produces):
+    def task_copy_to_root(depends_on: Path, produces: Path) -> None:
         """Copy a document to the root directory for easier retrieval."""
         shutil.copy(depends_on, produces)

### 2025-02-13 src/econ_spec_jel/config.py

@@ -1,7 +1,7 @@
 """Contains the general configuration of the project."""
 
-from typing import NamedTuple
 from pathlib import Path
+from typing import NamedTuple
 
 from pytask import DataCatalog
 
@@ -11,25 +11,39 @@ BLD = ROOT.joinpath("bld").resolve()
 DATA = ROOT.joinpath("data").resolve()
 DOCUMENTS = ROOT.joinpath("documents").resolve()
 
-DATACATALOGS = {"metadata": DataCatalog(name="metadata"),
-                "files": DataCatalog(name="files")}
+DATACATALOGS = {
+    "metadata": DataCatalog(name="metadata"),
+    "files": DataCatalog(name="files"),
+}
 
 MAX_DP_NUMBER = 17680
 
 
 class Scraper(NamedTuple):
-    """ NamedTuple to store the base urls and the dp number of a discussion paper."""
+    """NamedTuple to store the base urls and the dp number of a discussion paper."""
+
     metadata_base_url: str
     file_base_url: str
     dp_number: int
+
 
 SCRAPERS = [
     Scraper(
         metadata_base_url="https://www.iza.org/publications/dp/",
         file_base_url="https://docs.iza.org/dp",
-        dp_number=num
-        )
-    for num in range(1, MAX_DP_NUMBER+1)
-    ]
+        dp_number=num,
+    )
+    for num in range(1, MAX_DP_NUMBER + 1)
+]
 
-__all__ = ["BLD", "DATA", "DATACATALOGS", "DOCUMENTS", "MAX_DP_NUMBER", "ROOT", "Scraper", "SCRAPERS", "SRC"]
+__all__ = [
+    "BLD",
+    "DATA",
+    "DATACATALOGS",
+    "DOCUMENTS",
+    "MAX_DP_NUMBER",
+    "ROOT",
+    "SCRAPERS",
+    "SRC",
+    "Scraper",
+]

### 2025-02-13 src/econ_spec_jel/data_scraping/__init__.py

@@ -1,0 +1,1 @@
+"""Contains the namespace of data_scraping."""

### 2025-02-13 src/econ_spec_jel/data_scraping/helper.py

@@ -1,50 +1,56 @@
-from bs4 import BeautifulSoup
-import requests
+"""Helper module for data scraping."""
 
-def extract_metadata(response: requests.Response, dp_number: int) -> dict:
-    """ Extract metadata from the response of a discussion paper page.
+from typing import Any
+
+import requests
+from bs4 import BeautifulSoup
+
+
+def extract_metadata(response: requests.Response, dp_number: int) -> dict[str, Any]:
+    """Extract metadata from the response of a discussion paper page.
 
     Args:
         response (requests.Response): Response object of the discussion paper page.
         dp_number (int): Discussion paper number.
 
-    Returns:
+    Returns
+    -------
         dict: Metadata.
     """
     soup = BeautifulSoup(response.text, "html.parser")
 
-    title = soup.select_one('meta[property="og:title"]')['content'].strip()
-    pub_date = soup.select_one('div.publications-header p').text.strip()
+    title = soup.select_one('meta[property="og:title"]')["content"].strip()
+    pub_date = soup.select_one("div.publications-header p").text.strip()
     pub_month, pub_year = pub_date.split(" ")
-    authors_information = soup.select('div.authors a')
+    authors_information = soup.select("div.authors a")
     author_names = [a.text.strip() for a in authors_information]
-    author_urls = [a['href'] for a in authors_information]
-    abstract = soup.select_one('div.element-copyexpandable p').text.strip()
+    author_urls = [a["href"] for a in authors_information]
+    abstract = soup.select_one("div.element-copyexpandable p").text.strip()
     keywords = [
         li.text.strip()
-        for h3 in soup.find_all('h3', class_='box-list-headline')
-        if 'Keywords' in h3.text
-        for li in h3.find_next('ul').find_all('li')
+        for h3 in soup.find_all("h3", class_="box-list-headline")
+        if "Keywords" in h3.text
+        for li in h3.find_next("ul").find_all("li")
     ]
     jel_codes = [
         li.text.strip()
-        for h3 in soup.find_all('h3', class_='box-list-headline')
-        if 'JEL Codes' in h3.text
-        for li in h3.find_next('ul').find_all('li')
+        for h3 in soup.find_all("h3", class_="box-list-headline")
+        if "JEL Codes" in h3.text
+        for li in h3.find_next("ul").find_all("li")
     ]
     try:
-        file_url = soup.select_one('a.download-link')['href']
+        file_url = soup.select_one("a.download-link")["href"]
     except TypeError:
[…]

### 2025-02-13 src/econ_spec_jel/data_scraping/task.py

@@ -1,56 +1,82 @@
-""" Task functions for data scraping. """
-# write a loop over the two tasks and pass the number to each of them?
-# how else could I achieve parallelization/be sure not to need to re-run them all?
-import pytask
-from pytask import task
-import requests
-from typing import Annotated
-from pathlib import Path
+"""Task functions for data scraping."""
 
-from econ_spec_jel.config import DATACATALOGS, Scraper, SCRAPERS
+from pathlib import Path
+from typing import Annotated
+from typing import Any
+
+import requests
+from pytask import task
+
+from econ_spec_jel.config import DATACATALOGS
+from econ_spec_jel.config import SCRAPERS
+from econ_spec_jel.config import Scraper
 from econ_spec_jel.data_scraping.helper import extract_metadata
 
 
-def _scrape_metadata(base_url: str, dp_number: int) -> dict:
+def _scrape_metadata(base_url: str, dp_number: int) -> dict[str, Any]:
     url = f"{base_url}{dp_number}"
-    response = requests.get(url)
-    if response.status_code != 200:
-        return {'dp_number': dp_number} | {k: None for k in ['title', 'author_names', 'author_urls', 'publication_date_month', 'publication_date_year', 'abstract', 'keywords', 'jel_codes', 'file_url']}
+    response = requests.get(url, timeout=10)
+    if response.status_code != 200:  # PLR2004
+        return {"dp_number": dp_number} | {
+            k: None
+            for k in [
+                "title",
+                "author_names",
+                "author_urls",
+                "publication_date_month",
+                "publication_date_year",
+                "abstract",
+                "keywords",
+                "jel_codes",
+                "file_url",
+            ]
+        }
     return extract_metadata(response, dp_number)
+
 
 def _download_file(base_url: str, dp_number: int) -> bytes:
     url = f"{base_url}{dp_number}.pdf"
-    response = requests.get(url)
+    response = requests.get(url, timeout=10)
     return response.content
 
 
 for scraper in SCRAPERS:
     if not DATACATALOGS["metadata"][str(scraper.dp_number)].path.is_file():
+
         @task(id=str(scraper.dp_number))
         def task_scrape_metadata(
-            scraper: Scraper=scraper,
-            ) -> Annotated[Path, DATACATALOGS["metadata"][str(scraper.dp_number)]]:
-            """ Scrape metadata from discussion paper page.
+            scraper: Scraper = scraper,
+        ) -> Annotated[Path, DATACATALOGS["metadata"][str(scraper.dp_number)]]:
[…]

### 2025-02-13 src/econ_spec_jel/data_scraping/helper.py
Add scraping of information of publication.

@@ -23,6 +23,7 @@ def extract_metadata(response: requests.Response, dp_number: int) -> dict[str, A
     pub_date = soup.select_one("div.publications-header p").text.strip()
     pub_month, pub_year = pub_date.split(" ")
     authors_information = soup.select("div.authors a")
+    published = soup.select_one("div.publications-header p").text.strip()
     author_names = [a.text.strip() for a in authors_information]
     author_urls = [a["href"] for a in authors_information]
     abstract = soup.select_one("div.element-copyexpandable p").text.strip()
@@ -47,6 +48,7 @@ def extract_metadata(response: requests.Response, dp_number: int) -> dict[str, A
         "title": title,
         "author_names": author_names,
         "author_urls": author_urls,
+        "published": published,
         "publication_date_month": pub_month,
         "publication_date_year": pub_year,
         "abstract": abstract,

### 2025-02-13 src/econ_spec_jel/data_scraping/task.py

@@ -23,6 +23,7 @@ def _scrape_metadata(base_url: str, dp_number: int) -> dict[str, Any]:
                 "title",
                 "author_names",
                 "author_urls",
+                "published",
                 "publication_date_month",
                 "publication_date_year",
                 "abstract",

### 2025-02-13 src/econ_spec_jel/data_preparation/__init__.py
Merge metadata.

@@ -1,0 +1,1 @@
+"""Contains the namespace of data_preparation."""

### 2025-02-13 src/econ_spec_jel/data_preparation/task.py

@@ -1,0 +1,35 @@
+"""Data preparation tasks."""
+
+from pathlib import Path
+from typing import Annotated
+
+import pandas as pd
+
+from econ_spec_jel.config import DATACATALOGS
+
+
+def _merge_metadata(
+    data_catalog: Annotated[Path, DATACATALOGS["metadata"]],
+) -> pd.DataFrame:
+    data = [
+        DATACATALOGS["metadata"][entry].load()
+        for entry in data_catalog._entries
+        if entry != "merged"
+    ]  # SLF001
+    merged_data = pd.DataFrame.from_records(data).sort_values(by="dp_number")
+    return merged_data.reset_index(drop=True)
+
+
+def task_merge_metadata(
+    data_catalog: Annotated[Path, DATACATALOGS["metadata"]],
+) -> Annotated[Path, DATACATALOGS["metadata"]["merged"]]:
+    """Merge metadata of all discussion papers.
+
+    Args:
+        data_catalog: DataCatalog with metadata of all discussion papers.
+
+    Returns
+    -------
+        pd.DataFrame: Merged metadata.
+    """
+    return _merge_metadata(data_catalog=data_catalog)

### 2025-02-13 documents/paper.tex
Document missing urls and discussion papers.

@@
 
+There are 6 digits in the current range without a corresponding discussion paper accessible.
+The numbers are (4953, 12119, 12315, 12577, 12664, 16580).
+The first discussion paper would have been from May 2010.
+The later five ones fall into the time span January 2019 to November 2023.
+
+There are 9 discussion paper entries without a corresponding discussion paper file accessible.
+The numbers are (9, 18, 21, 28, 33, 56, 173, 16801, 17143).
+The first nine fall into the range June 1998 to July 2000.
+The later two are from February and July 2024.
+
+\clearpage
+
 \end{document}

### 2025-02-13 src/econ_spec_jel/data_preparation/task.py
Putting task functions at the top of files.

@@ -8,18 +8,6 @@ import pandas as pd
 from econ_spec_jel.config import DATACATALOGS
 
 
-def _merge_metadata(
-    data_catalog: Annotated[Path, DATACATALOGS["metadata"]],
-) -> pd.DataFrame:
-    data = [
-        DATACATALOGS["metadata"][entry].load()
-        for entry in data_catalog._entries
-        if entry != "merged"
-    ]  # SLF001
-    merged_data = pd.DataFrame.from_records(data).sort_values(by="dp_number")
-    return merged_data.reset_index(drop=True)
-
-
 def task_merge_metadata(
     data_catalog: Annotated[Path, DATACATALOGS["metadata"]],
 ) -> Annotated[Path, DATACATALOGS["metadata"]["merged"]]:
@@ -33,3 +21,15 @@ def task_merge_metadata(
         pd.DataFrame: Merged metadata.
     """
     return _merge_metadata(data_catalog=data_catalog)
+
+
+def _merge_metadata(
+    data_catalog: Annotated[Path, DATACATALOGS["metadata"]],
+) -> pd.DataFrame:
+    data = [
+        DATACATALOGS["metadata"][entry].load()
+        for entry in data_catalog._entries
+        if entry != "merged"
+    ]  # SLF001
+    merged_data = pd.DataFrame.from_records(data).sort_values(by="dp_number")
+    return merged_data.reset_index(drop=True)

### 2025-02-13 src/econ_spec_jel/data_scraping/task.py

@@ -12,35 +12,6 @@ from econ_spec_jel.config import SCRAPERS
 from econ_spec_jel.config import Scraper
 from econ_spec_jel.data_scraping.helper import extract_metadata
 
-
-def _scrape_metadata(base_url: str, dp_number: int) -> dict[str, Any]:
-    url = f"{base_url}{dp_number}"
-    response = requests.get(url, timeout=10)
-    if response.status_code != 200:  # PLR2004
-        return {"dp_number": dp_number} | {
-            k: None
-            for k in [
-                "title",
-                "author_names",
-                "author_urls",
-                "published",
-                "publication_date_month",
-                "publication_date_year",
-                "abstract",
-                "keywords",
-                "jel_codes",
-                "file_url",
-            ]
-        }
-    return extract_metadata(response, dp_number)
-
-
-def _download_file(base_url: str, dp_number: int) -> bytes:
-    url = f"{base_url}{dp_number}.pdf"
-    response = requests.get(url, timeout=10)
-    return response.content
-
-
 for scraper in SCRAPERS:
     if not DATACATALOGS["metadata"][str(scraper.dp_number)].path.is_file():
 
@@ -81,3 +52,31 @@ for scraper in SCRAPERS:
             return _download_file(
                 base_url=scraper.file_base_url, dp_number=scraper.dp_number
             )
+
+
+def _scrape_metadata(base_url: str, dp_number: int) -> dict[str, Any]:
+    url = f"{base_url}{dp_number}"
+    response = requests.get(url, timeout=10)
+    if response.status_code != 200:  # PLR2004
+        return {"dp_number": dp_number} | {
+            k: None
+            for k in [
+                "title",
+                "author_names",
+                "author_urls",
+                "published",
+                "publication_date_month",
+                "publication_date_year",
+                "abstract",
+                "keywords",
+                "jel_codes",
+                "file_url",
+            ]
+        }
+    return extract_metadata(response, dp_number)
+
+
+def _download_file(base_url: str, dp_number: int) -> bytes:
+    url = f"{base_url}{dp_number}.pdf"
+    response = requests.get(url, timeout=10)
+    return response.content

### 2025-02-13 src/econ_spec_jel/config.py
pre-commit hooks.

@@ -1,7 +1,6 @@
 """Contains the general configuration of the project."""
 
 from pathlib import Path
-from typing import NamedTuple
 
 from pytask import DataCatalog
 
@@ -18,24 +17,6 @@ DATACATALOGS = {
 
 MAX_DP_NUMBER = 17680
 
-
-class Scraper(NamedTuple):
-    """NamedTuple to store the base urls and the dp number of a discussion paper."""
-
-    metadata_base_url: str
-    file_base_url: str
-    dp_number: int
-
-
-SCRAPERS = [
-    Scraper(
-        metadata_base_url="https://www.iza.org/publications/dp/",
-        file_base_url="https://docs.iza.org/dp",
-        dp_number=num,
-    )
-    for num in range(1, MAX_DP_NUMBER + 1)
-]
-
 __all__ = [
     "BLD",
     "DATA",
@@ -43,7 +24,5 @@ __all__ = [
     "DOCUMENTS",
     "MAX_DP_NUMBER",
     "ROOT",
-    "SCRAPERS",
     "SRC",
-    "Scraper",
 ]

### 2025-02-13 src/econ_spec_jel/data_preparation/task.py

@@ -23,6 +23,10 @@ def task_merge_metadata(
     return _merge_metadata(data_catalog=data_catalog)
 
 
+def task_data_cleaning() -> None:
+    """Clean the data."""
+
+
 def _merge_metadata(
     data_catalog: Annotated[Path, DATACATALOGS["metadata"]],
 ) -> pd.DataFrame:

### 2025-02-13 src/econ_spec_jel/data_scraping/task.py

@@ -3,14 +3,32 @@
 from pathlib import Path
 from typing import Annotated
 from typing import Any
+from typing import NamedTuple
 
 import requests
 from pytask import task
 
 from econ_spec_jel.config import DATACATALOGS
-from econ_spec_jel.config import SCRAPERS
-from econ_spec_jel.config import Scraper
+from econ_spec_jel.config import MAX_DP_NUMBER
 from econ_spec_jel.data_scraping.helper import extract_metadata
+
+
+class Scraper(NamedTuple):
+    """NamedTuple to store the base urls and the dp number of a discussion paper."""
+
+    metadata_base_url: str
+    file_base_url: str
+    dp_number: int
+
+
+SCRAPERS = [
+    Scraper(
+        metadata_base_url="https://www.iza.org/publications/dp/",
+        file_base_url="https://docs.iza.org/dp",
+        dp_number=num,
+    )
+    for num in range(1, MAX_DP_NUMBER + 1)
+]
 
 for scraper in SCRAPERS:
     if not DATACATALOGS["metadata"][str(scraper.dp_number)].path.is_file():

### 2025-02-14 src/econ_spec_jel/data_scraping/task.py
pre-commit hooks (again).

@@ -1,15 +1,12 @@
 """Task functions for data scraping."""
 
 from pathlib import Path
-from typing import Annotated
-from typing import Any
-from typing import NamedTuple
+from typing import Annotated, Any, NamedTuple
 
 import requests
 from pytask import task
 
-from econ_spec_jel.config import DATACATALOGS
-from econ_spec_jel.config import MAX_DP_NUMBER
+from econ_spec_jel.config import DATACATALOGS, MAX_DP_NUMBER
 from econ_spec_jel.data_scraping.helper import extract_metadata
 
 

### 2025-02-14 documents/task.py
Update to pixi environment.

@@ -6,9 +6,7 @@ from pathlib import Path
 import pytask
 from pytask_latex import compilation_steps as cs
 
-from econ_spec_jel.config import BLD
-from econ_spec_jel.config import DOCUMENTS
-from econ_spec_jel.config import ROOT
+from econ_spec_jel.config import BLD, DOCUMENTS, ROOT
 
 documents = ["paper", "presentation"]
 

### 2025-02-14 src/econ_spec_jel/config.py

@@ -1,8 +1,13 @@
 """Contains the general configuration of the project."""
 
 from pathlib import Path
-
+import pandas as pd
 from pytask import DataCatalog
+
+pd.set_option("mode.copy_on_write", True)
+pd.set_option("future.infer_string", True)
+pd.set_option("future.no_silent_downcasting", True)
+pd.set_option("plotting.backend", "plotly")
 
 SRC = Path(__file__).parent.resolve()
 ROOT = SRC.joinpath("..", "..").resolve()
@@ -15,7 +20,7 @@ DATACATALOGS = {
     "files": DataCatalog(name="files"),
 }
 
-MAX_DP_NUMBER = 17680
+MAX_DP_NUMBER = 17695
 
 __all__ = [
     "BLD",

### 2025-02-14 src/econ_spec_jel/data_scraping/helper.py
Refactor data scraping.

@@ -23,7 +23,10 @@ def extract_metadata(response: requests.Response, dp_number: int) -> dict[str, A
     pub_date = soup.select_one("div.publications-header p").text.strip()
     pub_month, pub_year = pub_date.split(" ")
     authors_information = soup.select("div.authors a")
-    published = soup.select_one("div.publications-header p").text.strip()
+    try:
+        published = soup.select_one("div.published p").text.strip()
+    except AttributeError:
+        published = None
     author_names = [a.text.strip() for a in authors_information]
     author_urls = [a["href"] for a in authors_information]
     abstract = soup.select_one("div.element-copyexpandable p").text.strip()

### 2025-02-14 src/econ_spec_jel/data_scraping/task.py

@@ -1,7 +1,7 @@
 """Task functions for data scraping."""
 
 from pathlib import Path
-from typing import Annotated, Any, NamedTuple
+from typing import Annotated, Any
 
 import requests
 from pytask import task
@@ -10,88 +10,83 @@ from econ_spec_jel.config import DATACATALOGS, MAX_DP_NUMBER
 from econ_spec_jel.data_scraping.helper import extract_metadata
 
 
-class Scraper(NamedTuple):
-    """NamedTuple to store the base urls and the dp number of a discussion paper."""
-
-    metadata_base_url: str
-    file_base_url: str
-    dp_number: int
+def _metadata_has_not_been_scraped(dp_number: int) -> bool:
+    return not DATACATALOGS["metadata"][f"{dp_number}"].path.is_file()
 
 
-SCRAPERS = [
-    Scraper(
-        metadata_base_url="https://www.iza.org/publications/dp/",
-        file_base_url="https://docs.iza.org/dp",
-        dp_number=num,
-    )
-    for num in range(1, MAX_DP_NUMBER + 1)
-]
+def _file_has_not_been_downloaded(dp_number: int) -> bool:
+    return not DATACATALOGS["files"][f"{dp_number}"].path.is_file()
 
-for scraper in SCRAPERS:
-    if not DATACATALOGS["metadata"][str(scraper.dp_number)].path.is_file():
 
-        @task(id=str(scraper.dp_number))
+for dp_number in range(1, MAX_DP_NUMBER + 1):
+    if _metadata_has_not_been_scraped(dp_number=dp_number):
+
+        @task(id=f"{dp_number}")
         def task_scrape_metadata(
-            scraper: Scraper = scraper,
-        ) -> Annotated[Path, DATACATALOGS["metadata"][str(scraper.dp_number)]]:
-            """Scrape metadata from discussion paper page.
+            dp_number: int = dp_number,
+        ) -> Annotated[
+            Path,
+            (
+                DATACATALOGS["metadata"][f"{dp_number}"],
+                DATACATALOGS["files"][f"{dp_number}"],
+            ),
+        ]:
+            """Scrape discussion paper metadata.
 
             Args:
-                scraper: Scraper object with metadata_base_url and dp_number attributes.
+                dp_number: Discussion paper number.
 
             Returns
             -------
                 dict: Metadata.
             """
-            return _scrape_metadata(
-                base_url=scraper.metadata_base_url, dp_number=scraper.dp_number
-            )
+            return _scrape_metadata(dp_number=dp_number)
 
[…]

### 2025-02-17 src/econ_spec_jel/__init__.py
Fixing pixi.

@@ -1,4 +1,4 @@
-"""Contains the main namespace of econ-spec-jel."""
+"""Contains the main namespace of econ_spec_jel."""
 
 # Import the version from _version.py which is dynamically created by setuptools-scm
 # when the project is installed with ``pip install -e .``. Do not put it into version

### 2025-02-17 src/econ_spec_jel/data_scraping/task.py
Fix scraping task produces.

@@ -24,13 +24,7 @@ for dp_number in range(1, MAX_DP_NUMBER + 1):
         @task(id=f"{dp_number}")
         def task_scrape_metadata(
             dp_number: int = dp_number,
-        ) -> Annotated[
-            Path,
-            (
-                DATACATALOGS["metadata"][f"{dp_number}"],
-                DATACATALOGS["files"][f"{dp_number}"],
-            ),
-        ]:
+        ) -> Annotated[Path, DATACATALOGS["metadata"][f"{dp_number}"]]:
             """Scrape discussion paper metadata.
 
             Args:

### 2025-02-17 src/econ_spec_jel/data_preparation/task.py
Data cleaning (mainly publication information).

@@ -4,6 +4,7 @@ from pathlib import Path
 from typing import Annotated
 
 import pandas as pd
+import regex as re
 
 from econ_spec_jel.config import DATACATALOGS
 
@@ -14,7 +15,7 @@ def task_merge_metadata(
     """Merge metadata of all discussion papers.
 
     Args:
-        data_catalog: DataCatalog with metadata of all discussion papers.
+        data_catalog (pytask.DataCatalog): DataCatalog containing metadata.
 
     Returns
     -------
@@ -23,8 +24,19 @@ def task_merge_metadata(
     return _merge_metadata(data_catalog=data_catalog)
 
 
-def task_data_cleaning() -> None:
-    """Clean the data."""
+def task_data_cleaning(
+    merged_metadata: Annotated[Path, DATACATALOGS["metadata"]["merged"]],
+) -> Annotated[Path, DATACATALOGS["metadata"]["cleaned"]]:
+    """Clean the metadata.
+
+    Args:
+        merged_metadata (pd.DataFrame): Merged metadata.
+
+    Returns
+    -------
+        pd.DataFrame: Cleaned metadata.
+    """
+    return _clean_metadata(merged_metadata)
 
 
 def _merge_metadata(
@@ -33,7 +45,110 @@ def _merge_metadata(
     data = [
         DATACATALOGS["metadata"][entry].load()
         for entry in data_catalog._entries
-        if entry != "merged"
+        if entry not in ("merged", "cleaned")
     ]  # SLF001
     merged_data = pd.DataFrame.from_records(data).sort_values(by="dp_number")
     return merged_data.reset_index(drop=True)
+
+
+def _clean_metadata(
+    merged_metadata: pd.DataFrame,
+) -> pd.DataFrame:
+    metadata_complete_title_file_url = merged_metadata.dropna(
+        subset=["title", "file_url"]
+    )
+    cleaned_data = pd.DataFrame()
+    cleaned_data["dp_number"] = metadata_complete_title_file_url["dp_number"].astype(
+        "uint16[pyarrow]"
+    )
+    cleaned_data["title"] = metadata_complete_title_file_url["title"].astype(
+        "string[pyarrow]"
+    )
+    cleaned_data["author_names"] = metadata_complete_title_file_url["author_names"]
+    cleaned_data["author_urls"] = metadata_complete_title_file_url["author_urls"]
[…]

### 2025-02-17 src/econ_spec_jel/data_analysis/__init__.py
Initial commit of analysis.

@@ -1,0 +1,1 @@
+"""Contains the namespace of data_analysis."""

### 2025-02-17 src/econ_spec_jel/data_analysis/task.py

@@ -1,0 +1,29 @@
+"""Data analysis tasks."""
+
+from pathlib import Path
+from typing import Annotated
+
+import pandas as pd
+
+from econ_spec_jel.config import DATACATALOGS
+
+
+def task_data_analysis(
+    df: Annotated[Path, DATACATALOGS["metadata"]["cleaned"]],
+) -> Annotated[Path, DATACATALOGS["metadata"]["analysis"]]:
+    """Metadata analysis.
+
+    Args:
+        df (pd.DataFrame): DataCatalog containing cleaned metadata.
+
+    Returns
+    -------
+        pd.DataFrame: Metadata analysis.
+    """
+    return _data_analysis(df=df)
+
+
+def _data_analysis(
+    df: pd.DataFrame,
+) -> pd.DataFrame:
+    return df

### 2025-02-17 src/econ_spec_jel/data_preparation/task.py

@@ -45,7 +45,7 @@ def _merge_metadata(
     data = [
         DATACATALOGS["metadata"][entry].load()
         for entry in data_catalog._entries
-        if entry not in ("merged", "cleaned")
+        if entry not in ("merged", "cleaned", "analysis")
     ]  # SLF001
     merged_data = pd.DataFrame.from_records(data).sort_values(by="dp_number")
     return merged_data.reset_index(drop=True)

### 2025-02-17 src/econ_spec_jel/config.py
Refactor data catalogs.

@@ -16,8 +16,11 @@ DATA = ROOT.joinpath("data").resolve()
 DOCUMENTS = ROOT.joinpath("documents").resolve()
 
 DATACATALOGS = {
-    "metadata": DataCatalog(name="metadata"),
-    "files": DataCatalog(name="files"),
+    "raw": {
+        "metadata": DataCatalog(name="metadata"),
+        "files": DataCatalog(name="files"),
+    },
+    "data": DataCatalog(name="data"),
 }
 
 MAX_DP_NUMBER = 17695

### 2025-02-17 src/econ_spec_jel/data_analysis/task.py

@@ -9,8 +9,8 @@ from econ_spec_jel.config import DATACATALOGS
 
 
 def task_data_analysis(
-    df: Annotated[Path, DATACATALOGS["metadata"]["cleaned"]],
-) -> Annotated[Path, DATACATALOGS["metadata"]["analysis"]]:
+    df: Annotated[Path, DATACATALOGS["data"]["cleaned"]],
+) -> Annotated[Path, DATACATALOGS["data"]["analysis"]]:
     """Metadata analysis.
 
     Args:

### 2025-02-17 src/econ_spec_jel/data_preparation/task.py

@@ -5,13 +5,14 @@ from typing import Annotated
 
 import pandas as pd
 import regex as re
+import pytask
 
 from econ_spec_jel.config import DATACATALOGS
 
 
 def task_merge_metadata(
-    data_catalog: Annotated[Path, DATACATALOGS["metadata"]],
-) -> Annotated[Path, DATACATALOGS["metadata"]["merged"]]:
+    data_catalog: Annotated[Path, DATACATALOGS["raw"]["metadata"]],
+) -> Annotated[Path, DATACATALOGS["data"]["merged"]]:
     """Merge metadata of all discussion papers.
 
     Args:
@@ -25,8 +26,8 @@ def task_merge_metadata(
 
 
 def task_data_cleaning(
-    merged_metadata: Annotated[Path, DATACATALOGS["metadata"]["merged"]],
-) -> Annotated[Path, DATACATALOGS["metadata"]["cleaned"]]:
+    merged_metadata: Annotated[Path, DATACATALOGS["data"]["merged"]],
+) -> Annotated[Path, DATACATALOGS["data"]["cleaned"]]:
     """Clean the metadata.
 
     Args:
@@ -40,12 +41,10 @@ def task_data_cleaning(
 
 
 def _merge_metadata(
-    data_catalog: Annotated[Path, DATACATALOGS["metadata"]],
+    data_catalog: pytask.DataCatalog,
 ) -> pd.DataFrame:
     data = [
-        DATACATALOGS["metadata"][entry].load()
-        for entry in data_catalog._entries
-        if entry not in ("merged", "cleaned", "analysis")
+        data_catalog[entry].load() for entry in data_catalog._entries if entry.isdigit()
     ]  # SLF001
     merged_data = pd.DataFrame.from_records(data).sort_values(by="dp_number")
     return merged_data.reset_index(drop=True)

### 2025-02-17 src/econ_spec_jel/data_scraping/task.py

@@ -11,11 +11,11 @@ from econ_spec_jel.data_scraping.helper import extract_metadata
 
 
 def _metadata_has_not_been_scraped(dp_number: int) -> bool:
-    return not DATACATALOGS["metadata"][f"{dp_number}"].path.is_file()
+    return not DATACATALOGS["raw"]["metadata"][f"{dp_number}"].path.is_file()
 
 
 def _file_has_not_been_downloaded(dp_number: int) -> bool:
-    return not DATACATALOGS["files"][f"{dp_number}"].path.is_file()
+    return not DATACATALOGS["raw"]["files"][f"{dp_number}"].path.is_file()
 
 
 for dp_number in range(1, MAX_DP_NUMBER + 1):
@@ -24,7 +24,7 @@ for dp_number in range(1, MAX_DP_NUMBER + 1):
         @task(id=f"{dp_number}")
         def task_scrape_metadata(
             dp_number: int = dp_number,
-        ) -> Annotated[Path, DATACATALOGS["metadata"][f"{dp_number}"]]:
+        ) -> Annotated[Path, DATACATALOGS["raw"]["metadata"][f"{dp_number}"]]:
             """Scrape discussion paper metadata.
 
             Args:
@@ -41,7 +41,7 @@ for dp_number in range(1, MAX_DP_NUMBER + 1):
         @task(id=f"{dp_number}", after=f"task_scrape_metadata[{dp_number}]")
         def task_download_file(
             dp_number: int = dp_number,
-        ) -> Annotated[Path, DATACATALOGS["files"][f"{dp_number}"]]:
+        ) -> Annotated[Path, DATACATALOGS["raw"]["files"][f"{dp_number}"]]:
             """Download the discussion paper file.
 
             Args:

### 2025-02-18 documents/paper.tex
Update to pixi environment.
Initial commit of analysis.
Clean JEL codes for further analysis.

@@
 
+\section{Introduction}
+
+\section{Data}
+
+\subsection{Data Collection and Cleaning}
+
 There are 6 digits in the current range without a corresponding discussion paper accessible.
@@
 
+These 15 entries are excluded from the analysis.
+
+Dropped 18 entries, that have been superseded by another discussion paper.
+
+Dropped one discussion paper each with the JEL Codes "workfromhomecommutetimesallocationoftimesavingsCOVID19" (15870), and
+"labordisplacementenergytransitioncoalmines" (15581).
+Dropped 16 discussion papers due to completely missing JEL Codes, these were from March 2024 to February 2025.
+There were about 45 discussion papers that needed manual attention to fix the extracted JEL Codes.
+Likely the JEL Codes were extracted on IZA side via Optical Character Recognition (OCR) from the PDFs, leading to confusions of e.g. "I" and "1" or "O" and "0".
+
 \clearpage

### 2025-02-18 src/econ_spec_jel/data_preparation/task.py

@@ -74,21 +74,23 @@ def _clean_metadata(
         cleaned_data["other_publication_information"],
         cleaned_data["superseded"],
     ) = _clean_publication_information(cleaned_data["published_raw"])
-    cleaned_data["publication_date_month"] = metadata_complete_title_file_url[
+    cleaned_data["publication_month"] = metadata_complete_title_file_url[
         "publication_date_month"
     ].astype("string[pyarrow]")
-    cleaned_data["publication_date_year"] = metadata_complete_title_file_url[
+    cleaned_data["publication_year"] = metadata_complete_title_file_url[
         "publication_date_year"
     ].astype("uint16[pyarrow]")
     cleaned_data["abstract"] = metadata_complete_title_file_url["abstract"].astype(
         "string[pyarrow]"
     )
     cleaned_data["keywords"] = metadata_complete_title_file_url["keywords"]
-    cleaned_data["jel_codes"] = metadata_complete_title_file_url["jel_codes"]
+    cleaned_data["jel_codes"] = _clean_jel_codes(
+        metadata_complete_title_file_url["jel_codes"]
+    )
     cleaned_data["file_url"] = metadata_complete_title_file_url["file_url"].astype(
         "string[pyarrow]"
     )
-    return _not_superseded_dp(cleaned_data)
+    return _drop_dp_with_missing_data(cleaned_data)
 
 
 def _fuzzy_search_series_regex(
@@ -149,5 +151,106 @@ def _clean_publication_information(
     return published, forthcoming, other_publication_information, superseded
 
 
+def _clean_jel_codes(sr: pd.Series) -> pd.Series:
+    sr_manual_adapted = _apply_manual_adaptations(sr)
+    sr_capitalized = _capitalize_jel_codes(sr_manual_adapted)
+    sr_leading_i = _replace_leading_one_with_i(sr_capitalized)
+    sr_no_leading_zero = _drop_leading_zeros(sr_leading_i)
+    sr_proper_codes = _drop_improper_jel_codes(sr_no_leading_zero)
[…]

### 2025-02-19 src/econ_spec_jel/config.py
Fixing plotly.io and kaleido rendering problems.

@@ -3,6 +3,9 @@
 from pathlib import Path
 import pandas as pd
 from pytask import DataCatalog
+import plotly.io as pio
+
+pio.kaleido.scope.mathjax = None
 
 pd.set_option("mode.copy_on_write", True)
 pd.set_option("future.infer_string", True)

### 2025-02-19 src/econ_spec_jel/data_preparation/task.py
Refactor publication_year_month to datetime object.

@@ -9,6 +9,21 @@ import pytask
 
 from econ_spec_jel.config import DATACATALOGS
 
+MONTH_ORDER = [
+    "January",
+    "February",
+    "March",
+    "April",
+    "May",
+    "June",
+    "July",
+    "August",
+    "September",
+    "October",
+    "November",
+    "December",
+]
+
 
 def task_merge_metadata(
     data_catalog: Annotated[Path, DATACATALOGS["raw"]["metadata"]],
@@ -74,12 +89,10 @@ def _clean_metadata(
         cleaned_data["other_publication_information"],
         cleaned_data["superseded"],
     ) = _clean_publication_information(cleaned_data["published_raw"])
-    cleaned_data["publication_month"] = metadata_complete_title_file_url[
-        "publication_date_month"
-    ].astype("string[pyarrow]")
-    cleaned_data["publication_year"] = metadata_complete_title_file_url[
-        "publication_date_year"
-    ].astype("uint16[pyarrow]")
+    cleaned_data["publication_year_month"] = _create_publication_year_month(
+        metadata_complete_title_file_url["publication_date_year"],
+        metadata_complete_title_file_url["publication_date_month"],
+    )
     cleaned_data["abstract"] = metadata_complete_title_file_url["abstract"].astype(
         "string[pyarrow]"
     )
@@ -101,6 +114,15 @@ def _fuzzy_search_series_regex(
     return sr[
         sr.apply(lambda x: bool(re.search(re_pattern, str(x), flags=re.IGNORECASE)))
     ]
+
+
+def _create_publication_year_month(year: pd.Series, month: pd.Series) -> pd.Series:
+    month_numerical = month.map(
+        {month_name: i + 1 for i, month_name in enumerate(MONTH_ORDER)}
+    )
+    return pd.to_datetime(
+        year.astype("string[pyarrow]") + "-" + month_numerical.astype("string[pyarrow]")
+    )
 
 
 def _get_neither_published_nor_forthcoming(

### 2025-02-19 src/econ_spec_jel/data_analysis/task.py
Data analysis preparation and initital plot.

@@ -1,29 +1,0 @@
-"""Data analysis tasks."""
-
-from pathlib import Path
-from typing import Annotated
-
-import pandas as pd
-
-from econ_spec_jel.config import DATACATALOGS
-
-
-def task_data_analysis(
-    df: Annotated[Path, DATACATALOGS["data"]["cleaned"]],
-) -> Annotated[Path, DATACATALOGS["data"]["analysis"]]:
-    """Metadata analysis.
-
-    Args:
-        df (pd.DataFrame): DataCatalog containing cleaned metadata.
-
-    Returns
-    -------
-        pd.DataFrame: Metadata analysis.
-    """
-    return _data_analysis(df=df)
-
-
-def _data_analysis(
-    df: pd.DataFrame,
-) -> pd.DataFrame:
-    return df

### 2025-02-19 src/econ_spec_jel/data_analysis/task_plotting.py

@@ -1,0 +1,120 @@
+"""Data analysis tasks."""
+
+from pathlib import Path
+from typing import Annotated
+
+import pandas as pd
+import plotly.graph_objects as go
+import plotly.io as pio
+
+from econ_spec_jel.config import BLD, DATACATALOGS
+
+
+def task_plot_monthly_trends(
+    df: Annotated[Path, DATACATALOGS["data"]["analysis"]],
+) -> None:
+    """Plot 3-month trends of publications and average JEL codes per paper.
+
+    Args:
+        df (pd.DataFrame): Data prepared for analysis.
+    """
+    fig = _plot_monthly_trends(df=df)
+    pio.write_image(
+        fig, file=BLD / "figures" / "fig_counts_avg_jel_codes.pdf", format="pdf"
+    )
+
+
+def _plot_monthly_trends(df: pd.DataFrame) -> go.Figure:
+    # Aggregate data: Number of publications per year-month
+    pub_counts = (
+        df.groupby("publication_year_month").size().reset_index(name="pub_count")
+    )
+
+    # Aggregate data: Average JEL codes per paper per year-month
+    jel_avg = (
+        df.groupby("publication_year_month")["jel_codes_count"]
+        .mean()
+        .reset_index(name="avg_jel_per_pub")
+    )
+
+    # Merge both datasets
+    plot_data = pub_counts.merge(jel_avg, on="publication_year_month")
+
+    # Compute 6-month rolling average for smoothing
+    plot_data["pub_count_smooth"] = (
+        plot_data["pub_count"].rolling(6, center=False).mean()
+    )
+    plot_data["avg_jel_per_pub_smooth"] = (
+        plot_data["avg_jel_per_pub"].rolling(6, center=False).mean()
+    )
+    smooth_indices = plot_data.index[::6]
+
+    # Create figure
+    fig = go.Figure()
+
+    # Blurred original publication count line
+    fig.add_trace(
+        go.Scatter(
+            x=plot_data["publication_year_month"],
+            y=plot_data["pub_count"],
+            mode="lines",
+            name="Discussion Papers (monthly)",
+            line={"color": "blue", "width": 1, "dash": "dot"},  # Light & dotted line
+            opacity=0.3,  # Blur effect
+        )
+    )
+
+    # Smoothed publication count line (every third month only)
+    fig.add_trace(
+        go.Scatter(
+            x=plot_data.loc[smooth_indices, "publication_year_month"],
+            y=plot_data.loc[smooth_indices, "pub_count_smooth"],
[…]

### 2025-02-19 src/econ_spec_jel/data_analysis/task_preparation.py

@@ -1,0 +1,34 @@
+"""Data analysis tasks."""
+
+from pathlib import Path
+from typing import Annotated
+
+import pandas as pd
+
+from econ_spec_jel.config import DATACATALOGS
+
+
+def task_prepare_data_for_analysis(
+    df: Annotated[Path, DATACATALOGS["data"]["cleaned"]],
+) -> Annotated[Path, DATACATALOGS["data"]["analysis"]]:
+    """Prepare data for analysis.
+
+    Args:
+        df (pd.DataFrame): DataCatalog containing cleaned metadata.
+
+    Returns
+    -------
+        pd.DataFrame: Metadata prepared for analysis.
+    """
+    return _prepare_data_for_analysis(df=df)
+
+
+def _prepare_data_for_analysis(
+    df: pd.DataFrame,
+) -> pd.DataFrame:
+    out = df[
+        (df["publication_year_month"].dt.year >= 2000)
+        & (df["publication_year_month"].dt.year < 2025)
+    ].copy()  # PLR2004
+    out["jel_codes_count"] = out["jel_codes"].apply(len)
+    return out

### 2025-02-21 documents/paper.tex
Data analysis preparation and initital plot.
Renaming and refactoring of structure.

@@
 
-\title{Reproduction and Expansion of \\"SPECIALIZATION TRENDS IN ECONOMICS RESEARCH" by Sebastian Galiani, Ramiro H. Gálvez, and Ian Nachman}
+\title{Research Proposal: Specialization Trends in Economics Research using JEL Codes \\
+Text Data in Economics by Aapo Stenhammar \\
+University of Bonn, Winter Term 2024/25}
 
-\author{Felix Schmitz}
+\author{Felix Schmitz\footnotemark[1]}
 
 \date{
-    {\bf Under construction}
-    \\[1ex]
     \today
@@
 
+\subsection{Data Collection and Cleaning}
+For the analysis only data between January 2000 and December 2024 was selected.
+This is mainly to ensure a sufficient number of observations per month (for the early years) and complete information on the discussion paper (for the most recent discussion papers).
+This reduces the sample size by 195 observations. The final number of discussion papers analysed is 17,449.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_counts_avg_jel_codes}
+    \caption{Discussion Papers \& Avg. JEL Codes per Paper Over Time}
+    \label{fig:monthly_counts_and_avg_jel_codes_per_paper}
+\end{figure}
+
 \clearpage

### 2025-02-21 src/econ_spec_jel/analysis/task_plotting.py

@@ -116,5 +116,6 @@ def _plot_monthly_trends(df: pd.DataFrame) -> go.Figure:
             "color": "red",
         },
         legend={"x": 0.05, "y": 0.95},
+        margin={"l": 20, "r": 20, "t": 20, "b": 20},
     )
     return fig

### 2025-02-21 src/econ_spec_jel/data_management/clean_helper.py

@@ -1,13 +1,7 @@
-"""Data preparation tasks."""
-
-from pathlib import Path
-from typing import Annotated
+"""Helper functions for cleaning the metadata of discussion papers."""
 
 import pandas as pd
 import regex as re
-import pytask
-
-from econ_spec_jel.config import DATACATALOGS
 
 MONTH_ORDER = [
     "January",
@@ -25,87 +19,6 @@ MONTH_ORDER = [
 ]
 
 
-def task_merge_metadata(
-    data_catalog: Annotated[Path, DATACATALOGS["raw"]["metadata"]],
-) -> Annotated[Path, DATACATALOGS["data"]["merged"]]:
-    """Merge metadata of all discussion papers.
-
-    Args:
-        data_catalog (pytask.DataCatalog): DataCatalog containing metadata.
-
-    Returns
-    -------
-        pd.DataFrame: Merged metadata.
-    """
-    return _merge_metadata(data_catalog=data_catalog)
-
-
-def task_data_cleaning(
-    merged_metadata: Annotated[Path, DATACATALOGS["data"]["merged"]],
-) -> Annotated[Path, DATACATALOGS["data"]["cleaned"]]:
-    """Clean the metadata.
-
-    Args:
-        merged_metadata (pd.DataFrame): Merged metadata.
-
-    Returns
-    -------
-        pd.DataFrame: Cleaned metadata.
-    """
-    return _clean_metadata(merged_metadata)
-
-
-def _merge_metadata(
-    data_catalog: pytask.DataCatalog,
-) -> pd.DataFrame:
-    data = [
-        data_catalog[entry].load() for entry in data_catalog._entries if entry.isdigit()
-    ]  # SLF001
-    merged_data = pd.DataFrame.from_records(data).sort_values(by="dp_number")
-    return merged_data.reset_index(drop=True)
-
-
-def _clean_metadata(
-    merged_metadata: pd.DataFrame,
-) -> pd.DataFrame:
-    metadata_complete_title_file_url = merged_metadata.dropna(
-        subset=["title", "file_url"]
-    )
-    cleaned_data = pd.DataFrame()
-    cleaned_data["dp_number"] = metadata_complete_title_file_url["dp_number"].astype(
-        "uint16[pyarrow]"
-    )
-    cleaned_data["title"] = metadata_complete_title_file_url["title"].astype(
-        "string[pyarrow]"
-    )
-    cleaned_data["author_names"] = metadata_complete_title_file_url["author_names"]
[…]

### 2025-02-21 src/econ_spec_jel/data_management/task_clean.py

@@ -1,0 +1,62 @@
+"""Data preparation tasks."""
+
+from pathlib import Path
+from typing import Annotated
+
+import pandas as pd
+
+from econ_spec_jel.config import DATACATALOGS
+from econ_spec_jel.data_management.clean_helper import (
+    clean_publication_information,
+    clean_jel_codes,
+    create_publication_year_month,
+    drop_dp_with_missing_data,
+)
+
+
+def task_data_cleaning(
+    merged_metadata: Annotated[Path, DATACATALOGS["data"]["merged"]],
+) -> Annotated[Path, DATACATALOGS["data"]["cleaned"]]:
+    """Clean the metadata.
+
+    Args:
+        merged_metadata (pd.DataFrame): Merged metadata.
+
+    Returns
+    -------
+        pd.DataFrame: Cleaned metadata.
+    """
+    metadata_complete_title_file_url = merged_metadata.dropna(
+        subset=["title", "file_url"]
+    )
+    return _clean_metadata(metadata_complete_title_file_url)
+
+
+def _clean_metadata(
+    merged_metadata: pd.DataFrame,
+) -> pd.DataFrame:
+    cleaned_data = pd.DataFrame()
+    cleaned_data["dp_number"] = merged_metadata["dp_number"].astype("uint16[pyarrow]")
+    cleaned_data["title"] = merged_metadata["title"].astype("string[pyarrow]")
+    cleaned_data["author_names"] = merged_metadata["author_names"]
+    cleaned_data["author_urls"] = merged_metadata["author_urls"]
+    cleaned_data["published_raw"] = merged_metadata["published"].astype(
+        "string[pyarrow]"
+    )
+
+    (
+        cleaned_data["published"],
+        cleaned_data["forthcoming"],
+        cleaned_data["other_publication_information"],
+        cleaned_data["superseded"],
+    ) = clean_publication_information(cleaned_data["published_raw"])
+
+    cleaned_data["publication_year_month"] = create_publication_year_month(
+        merged_metadata["publication_date_year"],
+        merged_metadata["publication_date_month"],
+    )
+    cleaned_data["abstract"] = merged_metadata["abstract"].astype("string[pyarrow]")
+    cleaned_data["keywords"] = merged_metadata["keywords"]
+    cleaned_data["jel_codes"] = clean_jel_codes(merged_metadata["jel_codes"])
+    cleaned_data["file_url"] = merged_metadata["file_url"].astype("string[pyarrow]")
+    return drop_dp_with_missing_data(cleaned_data)

### 2025-02-21 src/econ_spec_jel/data_management/task_merge.py

@@ -1,0 +1,34 @@
+"""Data preparation tasks."""
+
+from pathlib import Path
+from typing import Annotated
+
+import pandas as pd
+import pytask
+
+from econ_spec_jel.config import DATACATALOGS
+
+
+def task_merge_metadata(
+    data_catalog: Annotated[Path, DATACATALOGS["raw"]["metadata"]],
+) -> Annotated[Path, DATACATALOGS["data"]["merged"]]:
+    """Merge metadata of all discussion papers.
+
+    Args:
+        data_catalog (pytask.DataCatalog): DataCatalog containing metadata.
+
+    Returns
+    -------
+        pd.DataFrame: Merged metadata.
+    """
+    return _merge_metadata(data_catalog=data_catalog)
+
+
+def _merge_metadata(
+    data_catalog: pytask.DataCatalog,
+) -> pd.DataFrame:
+    data = [
+        data_catalog[entry].load() for entry in data_catalog._entries if entry.isdigit()
+    ]  # SLF001
+    merged_data = pd.DataFrame.from_records(data).sort_values(by="dp_number")
+    return merged_data.reset_index(drop=True)

### 2025-02-21 src/econ_spec_jel/scraping/task.py

@@ -7,7 +7,7 @@ import requests
 from pytask import task
 
 from econ_spec_jel.config import DATACATALOGS, MAX_DP_NUMBER
-from econ_spec_jel.data_scraping.helper import extract_metadata
+from econ_spec_jel.scraping.helper import extract_metadata
 
 
 def _metadata_has_not_been_scraped(dp_number: int) -> bool:

### 2025-02-22 src/econ_spec_jel/analysis/task_plotting.py
Adapt comments.

@@ -25,22 +25,20 @@ def task_plot_monthly_trends(
 
 
 def _plot_monthly_trends(df: pd.DataFrame) -> go.Figure:
-    # Aggregate data: Number of publications per year-month
+    # Number of publications per year-month
     pub_counts = (
         df.groupby("publication_year_month").size().reset_index(name="pub_count")
     )
 
-    # Aggregate data: Average JEL codes per paper per year-month
+    # Average JEL codes per paper per year-month
     jel_avg = (
         df.groupby("publication_year_month")["jel_codes_count"]
         .mean()
         .reset_index(name="avg_jel_per_pub")
     )
 
-    # Merge both datasets
     plot_data = pub_counts.merge(jel_avg, on="publication_year_month")
-
-    # Compute 6-month rolling average for smoothing
+    # 6-month rolling average for smoothing
     plot_data["pub_count_smooth"] = (
         plot_data["pub_count"].rolling(6, center=False).mean()
     )
@@ -49,22 +47,19 @@ def _plot_monthly_trends(df: pd.DataFrame) -> go.Figure:
     )
     smooth_indices = plot_data.index[::6]
 
-    # Create figure
     fig = go.Figure()
-
-    # Blurred original publication count line
+    # Monthly publication count line
     fig.add_trace(
         go.Scatter(
             x=plot_data["publication_year_month"],
             y=plot_data["pub_count"],
             mode="lines",
             name="Discussion Papers (monthly)",
-            line={"color": "blue", "width": 1, "dash": "dot"},  # Light & dotted line
-            opacity=0.3,  # Blur effect
+            line={"color": "blue", "width": 1, "dash": "dot"},
+            opacity=0.3,
         )
     )
-
-    # Smoothed publication count line (every third month only)
+    # Smoothed publication count line
     fig.add_trace(
         go.Scatter(
             x=plot_data.loc[smooth_indices, "publication_year_month"],
@@ -75,20 +70,19 @@ def _plot_monthly_trends(df: pd.DataFrame) -> go.Figure:
         )
     )
 
-    # Blurred original avg. JEL codes line
+    # Monthly avg. JEL codes line
     fig.add_trace(
         go.Scatter(
             x=plot_data["publication_year_month"],
             y=plot_data["avg_jel_per_pub"],
             mode="lines",
             name="Avg. JEL Codes per Discussion Paper (monthly)",
[…]

### 2025-02-22 src/econ_spec_jel/analysis/plotting_helper.py
Plotting.

@@ -1,0 +1,229 @@
+"""Helper functions for plotting the analysis results."""
+
+import pandas as pd
+import plotly.graph_objects as go
+import plotly.io as pio
+from plotly.subplots import make_subplots
+
+from econ_spec_jel.config import BLD
+
+
+def plot_author_trends(data: pd.DataFrame) -> None:
+    """Plot the trends of the average number of authors per discussion paper.
+
+    Args:
+        data (pd.DataFrame): The analysis data.
+    """
+    author_avg = (
+        data.groupby("publication_year_month")["authors_count"]
+        .mean()
+        .reset_index(name="avg_authors_per_pub")
+    )
+
+    author_avg["avg_authors_per_pub_smooth"] = (
+        author_avg["avg_authors_per_pub"].rolling(6, center=False).mean()
+    )
+    smooth_indices = author_avg.index[::6]
+
+    fig = go.Figure()
+    fig.add_trace(
+        go.Scatter(
+            x=author_avg["publication_year_month"],
+            y=author_avg["avg_authors_per_pub"],
+            mode="lines",
+            name="Avg. Authors per Discussion Paper (monthly)",
+            line={"color": "green", "width": 1, "dash": "dot"},
+            opacity=0.3,
+        )
+    )
+    fig.add_trace(
+        go.Scatter(
+            x=author_avg.loc[smooth_indices, "publication_year_month"],
+            y=author_avg.loc[smooth_indices, "avg_authors_per_pub_smooth"],
+            mode="lines+markers",
+            name="Avg. Authors per Discussion Paper (6-month avg.)",
+            line={"color": "green", "width": 2},
+        )
+    )
+    fig.update_layout(
+        xaxis={
+            "title": "Year-Month",
+            "tickmode": "array",
+            "tickvals": author_avg["publication_year_month"][::12],
+            "tickformat": "%Y",
+        },
+        yaxis={"title": "Avg. Authors per Discussion Paper", "color": "green"},
+        legend={"x": 0.05, "y": 0.95},
+        margin={"l": 20, "r": 20, "t": 20, "b": 20},
+        width=3 * 300,
+        height=2 * 300,
+    )
+    pio.write_image(fig, file=BLD / "figures" / "fig_author_trends.png")
+
+
+def plot_monthly_trends(data: pd.DataFrame) -> None:
+    """Plot the monthly trends of discussion papers and the average JEL codes per paper.
+
+    Args:
[…]

### 2025-02-22 src/econ_spec_jel/analysis/task_plotting.py

@@ -3,113 +3,22 @@
 from pathlib import Path
 from typing import Annotated
 
-import pandas as pd
-import plotly.graph_objects as go
-import plotly.io as pio
+from econ_spec_jel.config import DATACATALOGS
+from econ_spec_jel.analysis.plotting_helper import (
+    plot_author_trends,
+    plot_dp_counts,
+    plot_monthly_trends,
+)
 
-from econ_spec_jel.config import BLD, DATACATALOGS
 
-
-def task_plot_monthly_trends(
-    df: Annotated[Path, DATACATALOGS["data"]["analysis"]],
+def task_descriptive_plots(
+    data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
 ) -> None:
-    """Plot 3-month trends of publications and average JEL codes per paper.
+    """Create descriptive plots.
 
     Args:
-        df (pd.DataFrame): Data prepared for analysis.
+        data (Path): Path to the analysis data.
     """
-    fig = _plot_monthly_trends(df=df)
-    pio.write_image(
-        fig, file=BLD / "figures" / "fig_counts_avg_jel_codes.pdf", format="pdf"
-    )
-
-
-def _plot_monthly_trends(df: pd.DataFrame) -> go.Figure:
-    # Number of publications per year-month
-    pub_counts = (
-        df.groupby("publication_year_month").size().reset_index(name="pub_count")
-    )
-
-    # Average JEL codes per paper per year-month
-    jel_avg = (
-        df.groupby("publication_year_month")["jel_codes_count"]
-        .mean()
-        .reset_index(name="avg_jel_per_pub")
-    )
-
-    plot_data = pub_counts.merge(jel_avg, on="publication_year_month")
-    # 6-month rolling average for smoothing
-    plot_data["pub_count_smooth"] = (
-        plot_data["pub_count"].rolling(6, center=False).mean()
-    )
-    plot_data["avg_jel_per_pub_smooth"] = (
-        plot_data["avg_jel_per_pub"].rolling(6, center=False).mean()
-    )
-    smooth_indices = plot_data.index[::6]
-
-    fig = go.Figure()
-    # Monthly publication count line
-    fig.add_trace(
-        go.Scatter(
-            x=plot_data["publication_year_month"],
-            y=plot_data["pub_count"],
-            mode="lines",
-            name="Discussion Papers (monthly)",
-            line={"color": "blue", "width": 1, "dash": "dot"},
-            opacity=0.3,
-        )
-    )
[…]

### 2025-02-22 src/econ_spec_jel/config.py

@@ -3,9 +3,6 @@
 from pathlib import Path
 import pandas as pd
 from pytask import DataCatalog
-import plotly.io as pio
-
-pio.kaleido.scope.mathjax = None
 
 pd.set_option("mode.copy_on_write", True)
 pd.set_option("future.infer_string", True)

### 2025-02-22 src/econ_spec_jel/data_management/task_preparation.py

@@ -31,4 +31,5 @@ def _prepare_data_for_analysis(
         & (df["publication_year_month"].dt.year < 2025)
     ].copy()  # PLR2004
     out["jel_codes_count"] = out["jel_codes"].apply(len)
+    out["authors_count"] = out["author_names"].apply(len)
     return out

### 2025-02-22 src/econ_spec_jel/analysis/plotting_helper.py
Add author timeseries plots.

@@ -14,14 +14,18 @@ def plot_author_trends(data: pd.DataFrame) -> None:
     Args:
         data (pd.DataFrame): The analysis data.
     """
-    author_avg = (
-        data.groupby("publication_year_month")["authors_count"]
-        .mean()
-        .reset_index(name="avg_authors_per_pub")
+    author_avg = _create_avg_and_smoothed_data(
+        data=data,
+        grouping="publication_year_month",
+        relevant_metric="authors_count",
+        avg_name="avg_authors_per_pub",
     )
-
-    author_avg["avg_authors_per_pub_smooth"] = (
-        author_avg["avg_authors_per_pub"].rolling(6, center=False).mean()
+    author_trend = _determine_author_trend(data=data)
+    author_trend["new_authors_smooth"] = (
+        author_trend["new_authors"].rolling(6, center=False).mean()
+    )
+    author_trend["returning_authors_smooth"] = (
+        author_trend["returning_authors"].rolling(6, center=False).mean()
     )
     smooth_indices = author_avg.index[::6]
 
@@ -45,6 +49,49 @@ def plot_author_trends(data: pd.DataFrame) -> None:
             line={"color": "green", "width": 2},
         )
     )
+    fig.add_trace(
+        go.Scatter(
+            x=author_trend["publication_year_month"],
+            y=author_trend["new_authors"],
+            mode="lines",
+            name="New Authors (monthly)",
+            line={"color": "red", "width": 1, "dash": "dot"},
+            opacity=0.3,
+            yaxis="y2",
+        )
+    )
+    fig.add_trace(
+        go.Scatter(
+            x=author_trend.loc[smooth_indices, "publication_year_month"],
+            y=author_trend.loc[smooth_indices, "new_authors_smooth"],
+            mode="lines+markers",
+            name="New Authors (6-month avg.)",
+            line={"color": "red", "width": 2},
+            yaxis="y2",
+        )
+    )
+    fig.add_trace(
+        go.Scatter(
+            x=author_trend["publication_year_month"],
+            y=author_trend["returning_authors"],
+            mode="lines",
+            name="Returning Authors (monthly)",
+            line={"color": "blue", "width": 1, "dash": "dot"},
+            opacity=0.3,
+            yaxis="y2",
+        )
+    )
+    fig.add_trace(
+        go.Scatter(
[…]

### 2025-02-25 src/econ_spec_jel/analysis/plotting_helper.py
Refactor author trend calculations.

@@ -20,13 +20,20 @@ def plot_author_trends(data: pd.DataFrame) -> None:
         relevant_metric="authors_count",
         avg_name="avg_authors_per_pub",
     )
-    author_trend = _determine_author_trend(data=data)
-    author_trend["new_authors_smooth"] = (
-        author_trend["new_authors"].rolling(6, center=False).mean()
+
+    new_authors = _create_avg_and_smoothed_data(
+        data=data,
+        grouping="publication_year_month",
+        relevant_metric="authors_new",
+        avg_name="new_authors",
     )
-    author_trend["returning_authors_smooth"] = (
-        author_trend["returning_authors"].rolling(6, center=False).mean()
+    returning_authors = _create_avg_and_smoothed_data(
+        data=data,
+        grouping="publication_year_month",
+        relevant_metric="authors_returning",
+        avg_name="returning_authors",
     )
+
     smooth_indices = author_avg.index[::6]
 
     fig = go.Figure()
@@ -51,44 +58,40 @@ def plot_author_trends(data: pd.DataFrame) -> None:
     )
     fig.add_trace(
         go.Scatter(
-            x=author_trend["publication_year_month"],
-            y=author_trend["new_authors"],
+            x=new_authors["publication_year_month"],
+            y=new_authors["new_authors"],
             mode="lines",
-            name="New Authors (monthly)",
+            name="Avg. New Authors per Discussion Paper (monthly)",
             line={"color": "red", "width": 1, "dash": "dot"},
             opacity=0.3,
-            yaxis="y2",
         )
     )
     fig.add_trace(
         go.Scatter(
-            x=author_trend.loc[smooth_indices, "publication_year_month"],
-            y=author_trend.loc[smooth_indices, "new_authors_smooth"],
+            x=new_authors.loc[smooth_indices, "publication_year_month"],
+            y=new_authors.loc[smooth_indices, "new_authors_smooth"],
             mode="lines+markers",
-            name="New Authors (6-month avg.)",
+            name="Avg. New Authors per Discussion Paper (6-month avg.)",
             line={"color": "red", "width": 2},
-            yaxis="y2",
         )
     )
     fig.add_trace(
         go.Scatter(
-            x=author_trend["publication_year_month"],
-            y=author_trend["returning_authors"],
+            x=returning_authors["publication_year_month"],
+            y=returning_authors["returning_authors"],
             mode="lines",
[…]

### 2025-02-25 src/econ_spec_jel/data_management/task_preparation.py

@@ -26,10 +26,81 @@ def task_prepare_data_for_analysis(
 def _prepare_data_for_analysis(
     df: pd.DataFrame,
 ) -> pd.DataFrame:
-    out = df[
-        (df["publication_year_month"].dt.year >= 2000)
-        & (df["publication_year_month"].dt.year < 2025)
-    ].copy()  # PLR2004
+    out = (
+        df[
+            (df["publication_year_month"].dt.year >= 2000)
+            & (df["publication_year_month"].dt.year < 2025)
+        ]
+        .copy()
+        .reset_index(drop=True)
+    )  # PLR2004
     out["jel_codes_count"] = out["jel_codes"].apply(len)
     out["authors_count"] = out["author_names"].apply(len)
+    out["authors_new"], out["authors_returning"] = (
+        _count_initial_and_returning_publication(out)
+    )
     return out
+
+
+def _count_initial_and_returning_publication(data: pd.DataFrame) -> tuple[pd.Series]:
+    """Count the number of initial and returning authors for each discussion paper."""
+    authors_with_dp = _get_authors_with_dp(data)
+
+    min_indices = _first_publication_indices(authors_with_dp)
+    authors_initial_publication = authors_with_dp.loc[min_indices].reset_index(
+        drop=True
+    )
+    authors_returning_publications = authors_with_dp.loc[
+        ~authors_with_dp.index.isin(min_indices)
+    ].reset_index(drop=True)
+
+    dp_initial_publication = _get_dp_initial_publication(
+        authors_initial_publication, "num_initial_publication"
+    )
+    dp_returning_publication = _get_dp_initial_publication(
+        authors_returning_publications, "num_returning_publication"
+    )
+
+    return _initial_and_returning_publications(
+        data, dp_initial_publication, dp_returning_publication
+    )
+
+
+def _get_authors_with_dp(data: pd.DataFrame) -> pd.Series:
+    authors = data.explode("author_names")[
+        ["author_names", "publication_year_month", "dp_number"]
+    ]
+    authors_sorted = authors.sort_values(
+        ["author_names", "publication_year_month", "dp_number"]
+    )
+    authors_sorted.columns = ["author_name", "publication_year_month", "dp_number"]
+    return authors_sorted.drop_duplicates().dropna().reset_index(drop=True)
+
+
+def _first_publication_indices(data: pd.DataFrame) -> pd.Series:
[…]

### 2025-02-25 src/econ_spec_jel/analysis/task_network.py
Network creation.

@@ -1,0 +1,55 @@
+"""Task to create network of JEL codes."""
+
+import networkx as nx
+
+from typing import Annotated
+from pathlib import Path
+import pandas as pd
+
+from econ_spec_jel.config import DATACATALOGS
+
+
+def task_create_network(
+    data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
+) -> None:
+    """Create network of JEL codes."""
+    co_occurrence = _create_co_occurrence_df(data)
+    _create_graph_network(co_occurrence)
+
+
+def _create_co_occurrence_df(data: pd.DataFrame) -> pd.DataFrame:
+    exploded = data.explode("jel_codes")
+    merged = exploded.merge(exploded, on="dp_number")
+    co_occurrence_df = (
+        merged[merged["jel_codes_x"] != merged["jel_codes_y"]]
+        .groupby(["jel_codes_x", "jel_codes_y"])
+        .size()
+        .reset_index(name="co_occurrence")
+    )
+    return co_occurrence_df.pivot_table(
+        index="jel_codes_x", columns="jel_codes_y", values="co_occurrence", fill_value=0
+    )
+
+
+def _create_graph_network(co_occurrence_matrix: pd.DataFrame) -> nx.Graph:
+    graph = nx.Graph()
+    edges = _create_edges(co_occurrence_matrix)
+    graph.add_weighted_edges_from(edges)
+    return graph
+
+
+def _create_edges(co_occurrence_matrix: pd.DataFrame) -> list:
+    edges = co_occurrence_matrix.reset_index().melt(
+        id_vars="jel_codes_x", var_name="jel_codes_y", value_name="weight"
+    )
+    non_zero_and_non_looping_edges = edges[
+        (edges["weight"] > 0) & (edges["jel_codes_x"] != edges["jel_codes_y"])
+    ]
+    non_zero_and_non_looping_edges["weight"] = (
+        non_zero_and_non_looping_edges["weight"] * 10
+    )
+    return (
+        non_zero_and_non_looping_edges[["jel_codes_x", "jel_codes_y", "weight"]]
+        .to_records(index=False)
+        .tolist()
+    )

### 2025-02-25 src/econ_spec_jel/data_management/clean_helper.py
Manually remove superseded discussion paper.

@@ -219,11 +219,16 @@ def drop_dp_with_missing_data(df: pd.DataFrame) -> pd.DataFrame:
 
     """
     df_unique_dp = _not_superseded_dp(df)
-    return _non_empty_jel(df_unique_dp)
+    df_manually_cleaned = _manually_clean(df_unique_dp)
+    return _non_empty_jel(df_manually_cleaned)
 
 
 def _not_superseded_dp(df: pd.DataFrame) -> pd.DataFrame:
     return df.drop(df[df["superseded"].notna()].index)
+
+
+def _manually_clean(df: pd.DataFrame) -> pd.DataFrame:
+    return df.drop(df[df.dp_number == 15057].index)  # removed since superseded
 
 
 def _non_empty_jel(df: pd.DataFrame) -> pd.DataFrame:

### 2025-02-26 src/econ_spec_jel/analysis/plotting_helper.py
Refactor plotting and add jel_code specifi plots.

@@ -1,18 +1,19 @@
 """Helper functions for plotting the analysis results."""
 
+from pathlib import Path
 import pandas as pd
+import plotly.express as px
 import plotly.graph_objects as go
 import plotly.io as pio
 from plotly.subplots import make_subplots
 
-from econ_spec_jel.config import BLD
 
-
-def plot_author_trends(data: pd.DataFrame) -> None:
+def plot_author_trends(data: pd.DataFrame, produces: Path) -> None:
     """Plot the trends of the average number of authors per discussion paper.
 
     Args:
         data (pd.DataFrame): The analysis data.
+        produces (Path): The path to save the plot.
     """
     author_avg = _create_avg_and_smoothed_data(
         data=data,
@@ -96,26 +97,28 @@ def plot_author_trends(data: pd.DataFrame) -> None:
     )
 
     fig.update_layout(
+        template="plotly_white",
         xaxis={
             "title": "Year-Month",
             "tickmode": "array",
             "tickvals": author_avg["publication_year_month"][::12],
             "tickformat": "%Y",
         },
-        yaxis={"title": "Avg. Authors per Discussion Paper"},
+        yaxis={"title": "Number of Authors"},
         legend={"x": 0.05, "y": 0.95},
         margin={"l": 20, "r": 20, "t": 20, "b": 20},
         width=3 * 300,
         height=2 * 300,
     )
-    pio.write_image(fig, file=BLD / "figures" / "fig_author_trends.png")
+    pio.write_image(fig, file=produces)
 
 
-def plot_monthly_trends(data: pd.DataFrame) -> None:
+def plot_monthly_trends(data: pd.DataFrame, produces: Path) -> None:
     """Plot the monthly trends of discussion papers and the average JEL codes per paper.
 
     Args:
         data (pd.DataFrame): The analysis data.
+        produces (Path): The path to save the plot.
     """
     jel_avg = _create_avg_and_smoothed_data(
         data=data,
@@ -142,7 +145,7 @@ def plot_monthly_trends(data: pd.DataFrame) -> None:
             x=plot_data["publication_year_month"],
             y=plot_data["pub_count"],
             mode="lines",
-            name="Discussion Papers (monthly)",
+            name="Number of Discussion Papers (monthly)",
             line={"color": "blue", "width": 1, "dash": "dot"},
             opacity=0.3,
         )
[…]

### 2025-02-26 src/econ_spec_jel/analysis/task_plotting.py

@@ -1,24 +1,77 @@
 """Data analysis tasks."""
 
+import pandas as pd
+import pathlib
 from pathlib import Path
 from typing import Annotated
+from pytask import task
 
-from econ_spec_jel.config import DATACATALOGS
-from econ_spec_jel.analysis.plotting_helper import (
-    plot_author_trends,
-    plot_dp_counts,
-    plot_monthly_trends,
-)
+from econ_spec_jel.config import DATACATALOGS, FIGURES
+from econ_spec_jel.analysis import plotting_helper
 
 
-def task_descriptive_plots(
-    data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
+def _fail_if_wrong_instance(input_, instance_type) -> None:  # ANN001
+    if not isinstance(input_, instance_type):
+        msg = (
+            f"Expected {input_.__name__} to be of "
+            f"type {instance_type}, got {type(input_)} instead."
+        )
+        raise TypeError(msg)
+
+
+def _fail_if_invalid_id(id_: str) -> None:
+    if id_ not in dir(plotting_helper):
+        msg = (
+            f"Expected {id_} to be a valid function in "
+            f"econ_spec_jel.analysis.plotting_helper."
+        )
+        raise ValueError(msg)
+
+
+def _error_handling(
+    data: pd.DataFrame, produces: Path, number_of_codes: int, id_: str
 ) -> None:
-    """Create descriptive plots.
+    for input_, type_ in zip(
+        [data, produces, number_of_codes, id_],
+        [pd.core.frame.DataFrame, pathlib.PosixPath, int, str],
+        strict=False,
+    ):
+        _fail_if_wrong_instance(input_, type_)
+    _fail_if_invalid_id(id_)
 
-    Args:
-        data (Path): Path to the analysis data.
-    """
-    plot_author_trends(data=data)
-    plot_dp_counts(data=data)
-    plot_monthly_trends(data=data)
+
+ID_KWARGS = {
+    "plot_author_trends": {"produces": FIGURES / "fig_author_trends.png"},
+    "plot_dp_counts": {"produces": FIGURES / "fig_counts_avg_jel_codes.png"},
+    "plot_monthly_trends": {"produces": FIGURES / "fig_dp_counts.png"},
+    "plot_most_common_jel_codes": {
+        "produces": FIGURES / "fig_top5overall_jel.png",
+        "number_of_codes": 5,
+    },
+    "plot_yearly_most_common_jel_codes": {
[…]

### 2025-02-26 src/econ_spec_jel/config.py

@@ -14,6 +14,7 @@ ROOT = SRC.joinpath("..", "..").resolve()
 BLD = ROOT.joinpath("bld").resolve()
 DATA = ROOT.joinpath("data").resolve()
 DOCUMENTS = ROOT.joinpath("documents").resolve()
+FIGURES = BLD.joinpath("figures").resolve()
 
 DATACATALOGS = {
     "raw": {
@@ -30,6 +31,7 @@ __all__ = [
     "DATA",
     "DATACATALOGS",
     "DOCUMENTS",
+    "FIGURES",
     "MAX_DP_NUMBER",
     "ROOT",
     "SRC",

### 2025-02-26 src/econ_spec_jel/data_management/task_preparation.py
Tokenize and stem abstracts.

@@ -1,4 +1,10 @@
-"""Data analysis tasks."""
+"""Data preparation tasks."""
+
+import nltk
+from nltk.tokenize import word_tokenize
+from nltk.corpus import stopwords
+from nltk.stem import PorterStemmer
+
 
 from pathlib import Path
 from typing import Annotated
@@ -39,7 +45,32 @@ def _prepare_data_for_analysis(
     out["authors_new"], out["authors_returning"] = (
         _count_initial_and_returning_publication(out)
     )
+    out["abstract_tokenized"] = _prepare_abstract(out)
     return out
+
+
+def _prepare_abstract(data: pd.DataFrame) -> pd.Series:
+    nltk.download("punkt")
+    nltk.download("stopwords")
+    tokens = _tokenize_abstract(data["abstract"])
+    return _stem_tokens(tokens)
+
+
+def _tokenize_abstract(sr: pd.Series) -> pd.Series:
+    stop_words = set(stopwords.words("english"))
+    cleaned_abstracts = sr.str.lower().str.replace(r"\W+", " ", regex=True)
+    # Tokenize and filter stop words
+    return pd.Series(
+        [
+            [word for word in word_tokenize(doc) if word not in stop_words]
+            for doc in cleaned_abstracts
+        ]
+    )
+
+
+def _stem_tokens(tokens: pd.Series) -> pd.Series:
+    stemmer = PorterStemmer()
+    return pd.Series([[stemmer.stem(token) for token in doc] for doc in tokens])
 
 
 def _count_initial_and_returning_publication(data: pd.DataFrame) -> tuple[pd.Series]:

### 2025-02-26 src/econ_spec_jel/config.py
LDA text method implementation for topic analysis.

@@ -22,9 +22,11 @@ DATACATALOGS = {
         "files": DataCatalog(name="files"),
     },
     "data": DataCatalog(name="data"),
+    "topic_model": DataCatalog(name="topic_model"),
 }
 
 MAX_DP_NUMBER = 17695
+NUM_TOPICS = 750
 
 __all__ = [
     "BLD",
@@ -33,6 +35,7 @@ __all__ = [
     "DOCUMENTS",
     "FIGURES",
     "MAX_DP_NUMBER",
+    "NUM_TOPICS",
     "ROOT",
     "SRC",
 ]

### 2025-02-26 src/econ_spec_jel/text_model/__init__.py

@@ -1,0 +1,1 @@
+"""Contains the namespace of text_model."""

### 2025-02-26 src/econ_spec_jel/text_model/task_model.py

@@ -1,0 +1,94 @@
+"""Tasks for text analysis via LDA topic modelling."""
+
+from econ_spec_jel.config import DATACATALOGS, NUM_TOPICS
+import pandas as pd
+from pathlib import Path
+from typing import Annotated
+from gensim.corpora import Dictionary
+from gensim.models import LdaModel
+
+
+def task_gensim_dictionary(
+    data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
+) -> Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]]:
+    """Create a gensim dictionary from the text data.
+
+    Args:
+        data (pd.DataFrame): Data with abstracts.
+
+    Returns
+    -------
+        Dictionary: Gensim dictionary.
+
+    """
+    return _get_gensim_dictionary(data["abstract_tokenized"])
+
+
+def task_corpus(
+    data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
+    gensim_dictionary: Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]],
+) -> Annotated[Path, DATACATALOGS["topic_model"]["corpus"]]:
+    """Create a corpus from the text data.
+
+    Args:
+        data (pd.DataFrame): Data with abstracts.
+        gensim_dictionary (Dictionary): Gensim dictionary.
+
+    Returns
+    -------
+        list: Corpus.
+
+    """
+    return [gensim_dictionary.doc2bow(doc) for doc in data["abstract_tokenized"]]
+
+
+def task_train_topic_model(
+    gensim_dictionary: Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]],
+    corpus: Annotated[Path, DATACATALOGS["topic_model"]["corpus"]],
+) -> Annotated[Path, DATACATALOGS["topic_model"]["lda"]]:
+    """Abstract text analysis via Latent Dirichlet Allocation (LDA) topic modelling.
+
+    Args:
+        gensim_dictionary(Dictionary): Dictionary of gensim containing tokens.
+        corpus (list): Corpus of tokenized words.
+
+    Returns
+    -------
+        LdaModel: Trained LDA model.
+    """
+    return _train_lda(
+        corpus=corpus,
+        gensim_dictionary=gensim_dictionary,
+        num_topics=NUM_TOPICS,
+        chunksize=5000,
+        passes=20,
+        iterations=400,
+    )  # PLR0913
+
+
[…]

### 2025-02-26 src/econ_spec_jel/data_management/task_clean.py
Allow for loading the merged data from bld.

@@ -15,34 +15,30 @@ from econ_spec_jel.data_management.clean_helper import (
 
 
 def task_data_cleaning(
-    merged_metadata: Annotated[Path, DATACATALOGS["data"]["merged"]],
+    merged_data: Annotated[Path, DATACATALOGS["data"]["merged"]],
 ) -> Annotated[Path, DATACATALOGS["data"]["cleaned"]]:
-    """Clean the metadata.
+    """Clean the data.
 
     Args:
-        merged_metadata (pd.DataFrame): Merged metadata.
+        merged_data (pd.DataFrame): Merged data.
 
     Returns
     -------
-        pd.DataFrame: Cleaned metadata.
+        pd.DataFrame: Cleaned data.
     """
-    metadata_complete_title_file_url = merged_metadata.dropna(
-        subset=["title", "file_url"]
-    )
-    return _clean_metadata(metadata_complete_title_file_url)
+    data_complete_title_file_url = merged_data.dropna(subset=["title", "file_url"])
+    return _clean_data(data_complete_title_file_url)
 
 
-def _clean_metadata(
-    merged_metadata: pd.DataFrame,
+def _clean_data(
+    merged_data: pd.DataFrame,
 ) -> pd.DataFrame:
     cleaned_data = pd.DataFrame()
-    cleaned_data["dp_number"] = merged_metadata["dp_number"].astype("uint16[pyarrow]")
-    cleaned_data["title"] = merged_metadata["title"].astype("string[pyarrow]")
-    cleaned_data["author_names"] = merged_metadata["author_names"]
-    cleaned_data["author_urls"] = merged_metadata["author_urls"]
-    cleaned_data["published_raw"] = merged_metadata["published"].astype(
-        "string[pyarrow]"
-    )
+    cleaned_data["dp_number"] = merged_data["dp_number"].astype("uint16[pyarrow]")
+    cleaned_data["title"] = merged_data["title"].astype("string[pyarrow]")
+    cleaned_data["author_names"] = merged_data["author_names"]
+    cleaned_data["author_urls"] = merged_data["author_urls"]
+    cleaned_data["published_raw"] = merged_data["published"].astype("string[pyarrow]")
 
     (
         cleaned_data["published"],
@@ -52,11 +48,11 @@ def _clean_metadata(
     ) = clean_publication_information(cleaned_data["published_raw"])
 
     cleaned_data["publication_year_month"] = create_publication_year_month(
[…]

### 2025-02-26 src/econ_spec_jel/data_management/task_merge.py

@@ -6,25 +6,56 @@ from typing import Annotated
 import pandas as pd
 import pytask
 
-from econ_spec_jel.config import DATACATALOGS
+from econ_spec_jel.config import BLD, DATACATALOGS
 
 
-def task_merge_metadata(
-    data_catalog: Annotated[Path, DATACATALOGS["raw"]["metadata"]],
-) -> Annotated[Path, DATACATALOGS["data"]["merged"]]:
-    """Merge metadata of all discussion papers.
+if Path(BLD / "data" / "merged_data.pkl").exists():
+
+    def task_load_merged_data_from_bld(
+        merged_data_path: Annotated[Path, BLD / "data" / "merged_data.pkl"],
+    ) -> Annotated[Path, DATACATALOGS["data"]["merged"]]:
+        """Load merged data from the BLD folder.
+
+        Args:
+            merged_data_path (pd.DataFrame): Merged data.
+
+        Returns
+        -------
+            pd.DataFrame: Merged data.
+        """
+        return pd.read_pickle(merged_data_path)
+
+else:
+
+    def task_merge_data(
+        data_catalog: Annotated[Path, DATACATALOGS["raw"]["data"]],
+    ) -> Annotated[Path, DATACATALOGS["data"]["merged"]]:
+        """Merge data of all discussion papers.
+
+        Args:
+            data_catalog (pytask.DataCatalog): DataCatalog containing data.
+
+        Returns
+        -------
+            pd.DataFrame: Merged data.
+        """
+        return _merge_data(data_catalog=data_catalog)
+
+
+@pytask.mark.skip()
+def task_write_merged_data_to_bld(
+    merged_data: Annotated[Path, DATACATALOGS["data"]["merged"]],
+    out_path: Path = BLD / "data" / "merged_data.pkl",
+) -> None:
+    """Write merged data to the BLD folder.
 
     Args:
-        data_catalog (pytask.DataCatalog): DataCatalog containing metadata.
-
-    Returns
-    -------
-        pd.DataFrame: Merged metadata.
+        merged_data (pd.DataFrame): Merged data.
     """
-    return _merge_metadata(data_catalog=data_catalog)
+    merged_data.to_pickle(out_path)
 
 
-def _merge_metadata(
+def _merge_data(
     data_catalog: pytask.DataCatalog,
 ) -> pd.DataFrame:
     data = [

### 2025-02-27 src/econ_spec_jel/scraping/task.py
Pytask mark skip tasks for reproducing based on merged data from bld.

@@ -4,6 +4,7 @@ from pathlib import Path
 from typing import Annotated, Any
 
 import requests
+import pytask
 from pytask import task
 
 from econ_spec_jel.config import DATACATALOGS, MAX_DP_NUMBER
@@ -22,6 +23,7 @@ for dp_number in range(1, MAX_DP_NUMBER + 1):
     if _metadata_has_not_been_scraped(dp_number=dp_number):
 
         @task(id=f"{dp_number}")
+        @pytask.mark.skip()
         def task_scrape_metadata(
             dp_number: int = dp_number,
         ) -> Annotated[Path, DATACATALOGS["raw"]["metadata"][f"{dp_number}"]]:
@@ -39,6 +41,7 @@ for dp_number in range(1, MAX_DP_NUMBER + 1):
     if _file_has_not_been_downloaded(dp_number=dp_number):
 
         @task(id=f"{dp_number}", after=f"task_scrape_metadata[{dp_number}]")
+        @pytask.mark.skip()
         def task_download_file(
             dp_number: int = dp_number,
         ) -> Annotated[Path, DATACATALOGS["raw"]["files"][f"{dp_number}"]]:

### 2025-02-27 src/econ_spec_jel/text_model/task_model.py

@@ -6,8 +6,10 @@ from pathlib import Path
 from typing import Annotated
 from gensim.corpora import Dictionary
 from gensim.models import LdaModel
+import pytask
 
 
+@pytask.mark.skip()
 def task_gensim_dictionary(
     data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
 ) -> Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]]:
@@ -24,6 +26,7 @@ def task_gensim_dictionary(
     return _get_gensim_dictionary(data["abstract_tokenized"])
 
 
+@pytask.mark.skip()
 def task_corpus(
     data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
     gensim_dictionary: Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]],
@@ -42,6 +45,7 @@ def task_corpus(
     return [gensim_dictionary.doc2bow(doc) for doc in data["abstract_tokenized"]]
 
 
+@pytask.mark.skip()
 def task_train_topic_model(
     gensim_dictionary: Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]],
     corpus: Annotated[Path, DATACATALOGS["topic_model"]["corpus"]],

### 2025-02-27 documents/paper.tex
Add author timeseries plots.
Latex updates.

@@
 \usepackage{minibox}
-% \usepackage{pdf14} % Enable for Manuscriptcentral -- can't handle pdf 1.5
 % \usepackage{endfloat} % Enable to move tables / figures to the end. Useful for some
 % submissions.
+\usepackage[backend=bibtex, style=authoryear]{biblatex}
+
+\addbibresource{refs.bib}
 
@@
 \maketitle
-\footnotetext[1]{Email: \href{mailto:s87fschm@uni-bonn.de}{s87fschm@uni-bonn.de}}
-
-\begin{abstract}
-    Some abstract here.
-\end{abstract}
+\footnotetext[1]{Email: \href{mailto:s87fschm@uni-bonn.de}{s87fschm@uni-bonn.de}, Matriculation Number: 50173514}
 
@@
 \section{Introduction}
+This paper proposes research to investigates specialization trends within economics research, focusing on whether subfields are becoming more specialized over time.
+Understanding these trends is crucial as increasing specialization can influence knowledge dissemination, collaboration patterns, and the evolution of economic thought.
+
+To analyze specialization, a robust measure is required.
+The focus will be on leveraging Journal of Economic Literature (JEL) Codes, which are widely used by economists to classify research.
+These codes provide a structured way to categorize papers into subfields, allowing for a systematic analysis of specialization trends.
+The JEL Codes are assigned by authors or editors and are intended to reflect the main topics of a paper.
+While they do not cover the methods or data used in a paper, they provide a useful starting point for understanding the content of a paper, and hence of economic research in general.
+By examining the intersections of JEL Codes across papers, this study assesses the depth and diversity of specialization within economics research.
+
+This research also aims to validate the relevance of JEL Codes as a tool for analyzing specialization trends.
+Additionally, as an empirical application, the goal is to explores specialization dynamics within the field of Labor Economics, providing a focused example of how specialization may evolve over time.
+Through this approach, the study contributes to a better understanding of how economic research is structured and how subfields are connected or fragmented.
+The findings may have implications for how economists engage in interdisciplinary work and how academic communities organize knowledge.
+
+The following sections of this research proposal are structured as follows.
+First, the literature review provides an overview of the existing research on specialization trends in economics and the use of JEL Codes.
+Secondly, the data used for the analysis is described, including the data collection process and pre-processing steps.
+Descriptive statistics are then presented to provide an overview of the data.
+Next, the methodology is outlined, focusing on the use of topic modeling and network analysis to identify specialization trends.
+Finally, the conclusion summarizes the research objectives and the potential implications of the study.
+It also outlines the next steps for the analysis and the expected contributions of the research.
+
+\section{Literature Review}
+The literature on specialization trends in economics research is limited, with few studies focusing on the evolution of subfields within the discipline.
+However, there is a growing interest in understanding how knowledge is structured and how research communities are organized \cite{heikkila2022a}.
+investigates the use of JEL codes in economics research, and their development over time.
+The study finds that JEL codes are widely used to classify research in economics, and highlights the potential for "bibliometric and scientometric research" using these codes.
 
 \section{Data}
+For the sake of simplicity, the analysis of this paper are based on the IZA Discussion Paper Series.
+The working paper series is well established in the field of Labor Economics and was easily accessible via scraping.
+The series has been initialized in April 1998 and has been continuously published since then.
+The data includes the JEL Codes, the name of the authors, the abstract and the publication date of the discussion paper, among others.
 
-\subsection{Data Collection and Cleaning}
+For a future project, it would be interesting to extend the analysis to other working paper series or general economics journals.
+A dataset that might be easily available for analysis is a combination of Constellate, Semantic Scholar, and manually classified publications into subfields of economics research, as produced by \cite{galiani2023a}.
+This dataset would provide a more comprehensive view of specialization trends in economics research, and would allow to join the work with .
+
+\subsection{Data Collection}
+The data was scraped from the IZA website using the Python library \texttt{requests} and modules from the 'html-parser-library' \texttt{bs4.BeautifulSoup}.
+For each discussion paper, IZA has a webpage with metadata and a link to the PDF file.
+Both the metadata and the PDF file were downloaded and stored locally.
+The metadata for each discussion paper contains a title, a list of authors, the publication date (year and month), the list of assigned JEL Codes, a list of self-assigned keywords, and an abstract, among others.
+It was also possible to extract links provided for individual authors, which, if they exist, reference an IZA profile page with further information on the author, or a personal website.
+The PDF files were not used for the analysis so far.
+For this analysis, data has been collected up to February 2025 (discussion paper number 17695).
 
@@
 The later five ones fall into the time span January 2019 to November 2023.
+It is not possible to tell whether these discussion papers have been removed from the IZA website or whether they have never been published.
+
+\subsection{Data Pre-Processing}
+Initially, the metadata of the discussion papers was merged into a single dataset.
+This dataset was then cleaned and pre-processed to ensure consistency and quality for the analysis.
 
@@
 The later two are from February and July 2024.
+These 15 entries (missing discussion paper or missing file) are excluded from the analysis.
 
-These 15 entries are excluded from the analysis.
+Variables were cleaned and efficient data types were chosen.
+Text information regarding publication was extracted from the metadata where possible.
+This information was then categorized into 'published', 'forthcoming', 'other\_publication\_information', 'superseded'.
+This categorization used text methods on Levenshtein distance, fuzzy search, and string matching, to identify typos in the publication information.
+The raw publication information was retained in a separate column.
+Information on superseded discussion papers was used, to exclude these from the analysis.
+This allowed to exclude discussion papers, which otherwise would have been counted twice in the analysis.
+The combination of text methods to identify superseded discussion papers was combined with the publication date of the discussion paper for this step.
+This led to dropping 18 entries.
 
-Dropped 18 entries, that have been superseded by another discussion paper.
+The JEL Codes were cleaned with high precision.
+Deviations from the expected format of alphanumeric characters, beginning with a capital letter and followed by two digits, were manually corrected.
+There were about 45 discussion papers that needed manual attention to fix the extracted JEL Codes.
+In addition, the following algorithmic methods were used to clean the JEL Codes.
+The letters were converted to uppercase, leading '1' and '0' were replaced by 'I' and 'O', respectively.
+This confusion is likely due to the extraction of the JEL Codes via Optical Character Recognition (OCR) from the PDFs.
+After these thoroughly cleaning steps, JEL Codes not starting with a capital letter followed followed by at least one and at most two digits were dropped.
+Finally, JEL Codes with a single digit were converted to two digits by adding a trailing zero.
+Overall, one discussion paper each was dropped with the JEL Codes "workfromhomecommutetimesallocationoftimesavingsCOVID19" (15870), and
+"labordisplacementenergytransitioncoalmines" (15581).
+Further, 16 discussion papers were dropped due to completely missing JEL Codes. These were from March 2024 to February 2025.
 
-Dropped one discussion paper each with the JEL Codes "workfromhomecommutetimesallocationoftimesavingsCOVID19" (15870), and
-"labordisplacementenergytransitioncoalmines" (15581).
-Dropped 16 discussion papers due to completely missing JEL Codes, these were from March 2024 to February 2025.
-There were about 45 discussion papers that needed manual attention to fix the extracted JEL Codes.
-Likely the JEL Codes were extracted on IZA side via Optical Character Recognition (OCR) from the PDFs, leading to confusions of e.g. "I" and "1" or "O" and "0".
-
-\subsection{Data Collection and Cleaning}
-For the analysis only data between January 2000 and December 2024 was selected.
+In a last step, the dataset was limited to discussion papers with an inclusive publication date between January 2000 and December 2024.
 This is mainly to ensure a sufficient number of observations per month (for the early years) and complete information on the discussion paper (for the most recent discussion papers).
 This reduces the sample size by 195 observations. The final number of discussion papers analysed is 17,449.
+
+\subsection{Descriptive Statistics}
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_author_trends}
+    \caption{Average Number of Authors Over Time}
+    \label{fig:author_trends}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_dp_counts}
+    \caption{Number of Authors \& JEL Codes per Discussion Paper}
+    \label{fig:dp_counts}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_top5overall_jel}
+    \caption{Top 5 JEL Codes}
+    \label{fig:top5overall_jel}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_top3yearly_jel}
+    \caption{Top 3 JEL Codes yearly}
+    \label{fig:top3yearly_jel}
+\end{figure}
+
+\section{Methodology}
+
+\subsection{Topic Modeling}
+To analyze the specialization trends in economics research, topic modeling is used.
+Specifically, the Latent Dirichlet Allocation (LDA) algorithm is used to identify topics in the abstracts of the discussion papers \cite{blei2003}.
+To prepare the abstract data for the LDA model, the text is lowercased, tokenized, and stemmed.
+Numbers, punctuation, and words containing no alphanumeric characters are removed.
+Furthermore, stopwords are removed from the text.
+The model
+The number of topics is set to 750 and 20 passes through the corpus of tokenized and stemmed abstracts are performed during training of the model, similar to \cite{galiani2023a}.
+
+\subsection{Topic Modeling}
+
+To analyze specialization trends in economics research, topic modeling is employed to identify latent topics within the abstracts of discussion papers.
+Specifically, the Latent Dirichlet Allocation (LDA) generative model is used, which assumes that documents discussing similar topics tend to use similar groups of words \cite{blei2003}.
+LDA is particularly suitable for this analysis as it captures co-occurring word clusters, revealing thematic structures within the corpus.
+This approach models documents as random mixtures over latent topics, where each topic is characterized by a distribution over words.
+
+To prepare the abstract data for the LDA model, several preprocessing steps are conducted to enhance the quality of the input text.
+First, all text is converted to lowercase.
+Tokenization is then performed, splitting the text into individual words.
+Stemming is applied to reduce words to their root form, capturing the core meaning of words (e.g., replacing "running" with "run").
+Additionally, numbers, punctuation, and tokens without alphanumeric characters are removed.
+Stopwords, which are common function words with little semantic value, are also eliminated.
+These preprocessing steps collectively improve the quality of the tokenized text, ensuring a more effective training process for the LDA model.
+
+The LDA model is trained using \texttt{Gensim}'s implementation of the LDA training algorithm
+%\cite{rehurek2010}.
+The number of topics is set to 750, and the model is trained using 20 passes through the corpus.
+These settings are chosen to ensure sufficient topic granularity while maintaining interpretability, similar to the approach used in \cite{galiani2023a}.
+
+% is this true?
+The model uses a symmetric Dirichlet prior for the topic distribution per document ($\alpha$) and the word distribution per topic ($\eta$), reflecting the assumption that all topics and words are equally likely a priori.
+
+Two main outputs are obtained from the trained LDA model.
+First, the word distribution over topics, denoted as $w_{k,l}$, represents the probability of word $l$ in topic $k$, with $\sum_l w_{k,l} = 1$.
+This indicates the contribution of each word to the topic.
+Second, the topic distribution over documents, denoted as $t_{i,k}$, captures the presence of topic $k$ in document $i$, with $\sum_k t_{i,k} = 1$.
+This shows the relevance of each topic within the document.
+
+To ensure the meaningfulness of the detected topics, a filtering process is applied.
+Topics predominantly composed of non-thematic words (e.g., "online", "access", "appendix") or those lacking a clear thematic pattern are excluded.
+% After this filtering process, 695 meaningful topics are retained.
+An article $i$ is defined as "containing" topic $k$ if $t_{i,k}$ ranks among the top ten topics with the highest presence in that article.
+This approach ensures consistent topic assignment while maintaining interpretability.
+
+This topic modeling approach not only identifies prevalent themes but also captures the interconnectedness of topics within the field.
+It provides a robust foundation for analyzing specialization trends in economics research by revealing how topics cluster and evolve over time.
+The results from this robust methodology offer a comprehensive benchmark for understanding the dynamics of specialization within the field when using JEL codes.
+
+\subsection{JEL Code Analysis}
+
+\subsection{Overlapping Network Analysis}
+Initially, I wanted to use the Louvain algorithm to detect communities in the network.
+However, the algorithm only allows to non-overlapping communities, that is, a node can only belong to one community.
+This is not suitable for the analysis of specialization trends in economics research, as a JEL code can belong to multiple subfields.
+Rather the combination of JEL codes describes a subfield.
+Subsets of JEL codes describing a subfield might overlap with JEL codes describing another subfields.
+Hence, the goal is to use a different algorithm to detect overlapping communities in the network.
+
 \clearpage
 
+\section{Conclusion}
+In a next step, the self-assigned keywords of the discussion papers can be used for comparison with the results from the topic modeling and JEL Code analysis.
+
+\printbibliography
+
 \end{document}

### 2025-02-27 documents/task.py

@@ -8,9 +8,19 @@ from pytask_latex import compilation_steps as cs
 
 from econ_spec_jel.config import BLD, DOCUMENTS, ROOT
 
-documents = ["paper", "presentation"]
+figures = [
+    "fig_counts_avg_jel_codes.png",
+    "fig_author_trends.png",
+    "fig_dp_counts.png",
+    "fig_top5overall_jel.png",
+    "fig_top3yearly_jel.png",
+]
+DOCUMENTS_KWARGS = {
+    "paper": {"depends_on": [BLD / "figures" / figure for figure in figures]},
+    "presentation": {"depends_on": None},
+}
 
-for document in documents:
+for document, kwargs in DOCUMENTS_KWARGS.items():
 
     @pytask.mark.latex(
         script=DOCUMENTS / f"{document}.tex",
@@ -19,16 +29,16 @@ for document in documents:
             options=("--pdf", "--interaction=nonstopmode", "--synctex=1", "--cd"),
         ),
     )
-    @pytask.task(id=document)
-    def task_compile_document() -> None:
+    @pytask.task(id=document, kwargs=kwargs)
+    def task_compile_document(depends_on: None | list[Path]) -> None:
         """Compile the document specified in the latex decorator."""
 
-    kwargs = {
+    copy_to_root_kwargs = {
         "depends_on": BLD / "documents" / f"{document}.pdf",
         "produces": ROOT / f"{document}.pdf",
     }
 
-    @pytask.task(id=document, kwargs=kwargs)
+    @pytask.task(id=document, kwargs=copy_to_root_kwargs)
     def task_copy_to_root(depends_on: Path, produces: Path) -> None:
         """Copy a document to the root directory for easier retrieval."""
         shutil.copy(depends_on, produces)

### 2025-02-27 src/econ_spec_jel/analysis/task_descriptive_plots.py
Analysis and plotting.

@@ -42,7 +42,7 @@ def _error_handling(
 
 ID_KWARGS = {
     "plot_author_trends": {"produces": FIGURES / "fig_author_trends.png"},
-    "plot_dp_counts": {"produces": FIGURES / "fig_counts_avg_jel_codes.png"},
+    "plot_dp_counts": {"produces": FIGURES / "fig_counts_dp_jel_codes.png"},
     "plot_monthly_trends": {"produces": FIGURES / "fig_dp_counts.png"},
     "plot_most_common_jel_codes": {
         "produces": FIGURES / "fig_top5overall_jel.png",
@@ -57,13 +57,13 @@ ID_KWARGS = {
 for id_, kwargs_ in ID_KWARGS.items():
 
     @task(id=id_, kwargs=kwargs_)
-    def task_plot(
+    def task_(
         data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
         produces: Path,
         id_: str = id_,
         number_of_codes: int = 0,
     ) -> None:
-        """Create a plot.
+        """Create a plot based on the merged analysis data.
 
         Args:
             data (Path): Path to the analysis data.

### 2025-02-27 src/econ_spec_jel/analysis/task_topics_plot.py

@@ -1,0 +1,108 @@
+"""Data analysis tasks."""
+
+from pathlib import Path
+from typing import Annotated
+from pytask import task
+import pandas as pd
+
+from econ_spec_jel.config import DATACATALOGS, FIGURES
+from plotly.subplots import make_subplots
+import plotly.io as pio
+import plotly.graph_objects as go
+
+import seaborn as sns
+
+
+@task(kwargs={"produces": FIGURES / "fig_topic_trends.png"})
+def task_topic_trends(
+    plotting_data: Annotated[Path, DATACATALOGS["data"]["topics_plotting_data"]],
+    slopes: Annotated[Path, DATACATALOGS["data"]["slopes"]],
+    produces: Path,
+) -> None:
+    """Create a plot based on the topic data."""
+    _plot_topic_trends(plotting_data, slopes, produces)
+
+
+def _plot_topic_trends(
+    plotting_data: pd.DataFrame, slopes: pd.DataFrame, produces: Path
+) -> None:
+    top_up = slopes.head(5)["topic_number"].tolist()
+    top_down = slopes.tail(5)["topic_number"].tolist()
+
+    top_topics = top_up + top_down
+    relevant_topics = plotting_data[plotting_data["topic_number"].isin(top_topics)]
+
+    palette = sns.color_palette("husl", 5)
+    colors = [
+        f"rgba({int(r*255)}, {int(g*255)}, {int(b*255)}, 1)" for r, g, b in palette
+    ]
+
+    fig = make_subplots(
+        rows=5,
+        cols=2,
+        subplot_titles=[
+            str(
+                plotting_data.loc[
+                    plotting_data["topic_number"] == topic, "topic_words_8"
+                ].iloc[0]
+            )
+            for topic in top_topics
+        ],
+        shared_xaxes="columns",
+        shared_yaxes="all",
+        vertical_spacing=0.01,
+        horizontal_spacing=0.05,
+    )
+
+    for _i, (trend, topic_numbers) in enumerate(
+        zip(["Upward", "Downward"], [top_up, top_down], strict=False)
+    ):
+        col = 1 if trend == "Upward" else 2
+        for j, topic_number in enumerate(topic_numbers):
+            row = j + 1
+            df_topic = relevant_topics[relevant_topics["topic_number"] == topic_number]
+            plotting_data = (
+                df_topic.groupby("publication_year_month")["topic_present"]
+                .sum()
+                .reset_index()
+            )
[…]

### 2025-02-27 src/econ_spec_jel/data_management/task_preparation.py

@@ -58,7 +58,7 @@ def _prepare_abstract(data: pd.DataFrame) -> pd.Series:
 
 def _tokenize_abstract(sr: pd.Series) -> pd.Series:
     stop_words = set(stopwords.words("english"))
-    cleaned_abstracts = sr.str.lower().str.replace(r"\W+", " ", regex=True)
+    cleaned_abstracts = sr.str.lower().str.replace(r"[^a-zA-Z]+", " ", regex=True)
     # Tokenize and filter stop words
     return pd.Series(
         [

### 2025-02-27 src/econ_spec_jel/data_management/task_topics_preparation.py

@@ -1,0 +1,89 @@
+"""Module contains the task to prepare the topics data for the analysis."""
+
+from econ_spec_jel.config import DATACATALOGS
+from typing import Annotated
+from pathlib import Path
+import pandas as pd
+import numpy as np
+import statsmodels.api as sm
+
+
+def task_documents_topics_melted(
+    documents_topics: Annotated[Path, DATACATALOGS["data"]["documents_topics"]],
+) -> Annotated[Path, DATACATALOGS["data"]["documents_topics_melted"]]:
+    """Prepare the documents_topics data for the analysis."""
+    return _get_documents_topics_melted(documents_topics)
+
+
+def _get_documents_topics_melted(documents_topics: pd.DataFrame) -> pd.DataFrame:
+    """Efficiently melt and transform the documents_topics DataFrame."""
+    melted = pd.melt(
+        documents_topics,
+        id_vars=["dp_number", "title", "publication_year_month"],
+        value_vars=[f"top_{i:03d}" for i in range(750)],
+        var_name="topic_number",
+        value_name="topic_present",
+    )
+    melted["topic_number"] = melted["topic_number"].str[4:].astype(int)
+    return melted
+
+
+def task_slopes(
+    documents_topics_melted: Annotated[
+        Path, DATACATALOGS["data"]["documents_topics_melted"]
+    ],
+) -> Annotated[Path, DATACATALOGS["data"]["slopes"]]:
+    """Calculate the slopes of the topics."""
+    return _get_slopes(documents_topics_melted)
+
+
+def _get_slopes(data: pd.DataFrame) -> pd.DataFrame:
+    data["publication_ordinal"] = pd.to_datetime(data["publication_year_month"]).map(
+        pd.Timestamp.toordinal
+    )
+
+    slopes_df = (
+        data.groupby("topic_number").apply(_calculate_slope).reset_index(name="slope")
+    )
+    slopes_df = slopes_df.dropna()  # Drop NaNs resulting from insufficient data points
+    return slopes_df.sort_values(by="slope", ascending=False)
+
+
+def _calculate_slope(group: pd.DataFrame) -> float:
+    if group["topic_present"].notna().sum() > 1:
+        X = sm.add_constant(group["publication_ordinal"])
+        y = group["topic_present"].to_numpy()
+        model = sm.OLS(y, X).fit()
+        return model.params[1]
+    return np.nan
+
+
[…]

### 2025-02-27 src/econ_spec_jel/text_model/task_model.py

@@ -6,10 +6,9 @@ from pathlib import Path
 from typing import Annotated
 from gensim.corpora import Dictionary
 from gensim.models import LdaModel
-import pytask
 
 
-@pytask.mark.skip()
+# @pytask.mark.skip()
 def task_gensim_dictionary(
     data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
 ) -> Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]]:
@@ -26,7 +25,7 @@ def task_gensim_dictionary(
     return _get_gensim_dictionary(data["abstract_tokenized"])
 
 
-@pytask.mark.skip()
+# @pytask.mark.skip()
 def task_corpus(
     data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
     gensim_dictionary: Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]],
@@ -45,7 +44,7 @@ def task_corpus(
     return [gensim_dictionary.doc2bow(doc) for doc in data["abstract_tokenized"]]
 
 
-@pytask.mark.skip()
+# @pytask.mark.skip()
 def task_train_topic_model(
     gensim_dictionary: Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]],
     corpus: Annotated[Path, DATACATALOGS["topic_model"]["corpus"]],

### 2025-02-27 src/econ_spec_jel/text_model/task_topics.py

@@ -1,0 +1,72 @@
+"""Tasks for extracting topics from the trained LDA model."""
+
+from econ_spec_jel.config import DATACATALOGS, NUM_TOPICS
+import pandas as pd
+from pathlib import Path
+from typing import Annotated
+from gensim.models import LdaModel
+
+
+def task_extract_topics(
+    topic_model_data_catalog: Annotated[Path, DATACATALOGS["topic_model"]],
+) -> Annotated[Path, DATACATALOGS["data"]["topics"]]:
+    """Extract topics from the trained LDA model.
+
+    Args:
+        topic_model_data_catalog (Path): Path to the topic_model DataCatalog.
+
+    Returns
+    -------
+        (pd.DataFrame): Dataframe with the topics.
+    """
+    return _extract_topics(topic_model_data_catalog["lda"].load(), NUM_TOPICS)
+
+
+def task_extract_documents_topics_distribution(
+    topic_model_data_catalog: Annotated[Path, DATACATALOGS["topic_model"]],
+    data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
+) -> Annotated[Path, DATACATALOGS["data"]["documents_topics"]]:
+    """Extract distribution of documents to topics.
+
+    Args:
+        topic_model_data_catalog (Path): Path to the topic_model DataCatalog.
+        data (pd.DataFrame): Dataframe with the abstracts.
+
+    Returns
+    -------
+        (pd.DataFrame): Dataframe with the topics.
+    """
+    topic_model = topic_model_data_catalog["lda"].load()
+    corpus = topic_model_data_catalog["corpus"].load()
+    return _extract_documents_topics_distribution(data, topic_model, corpus)
+
+
+def _extract_topics(topic_model: LdaModel, num_topics: int) -> pd.DataFrame:
+    detected_topics = {}
+    for topic_num in range(num_topics):
+        detected_topics[topic_num] = [
+            f"{t} ({100 * p:.10f}%)" for t, p in topic_model.show_topic(topic_num, 20)
+        ]
+
+    detected_topics_df = pd.DataFrame(detected_topics).T
+    detected_topics_df.columns = [f"token_{e}" for e in range(20)]
+    return detected_topics_df
+
+
+def _extract_documents_topics_distribution(
+    data: pd.DataFrame, topic_model: LdaModel, corpus: list
+) -> pd.DataFrame:
+    documents_topics = []
+    for idx, dp in enumerate(data.to_dict(orient="records")):
[…]

### 2025-02-27 documents/paper.tex
Paper writing.

@@
 \footnotetext[1]{Email: \href{mailto:s87fschm@uni-bonn.de}{s87fschm@uni-bonn.de}, Matriculation Number: 50173514}
-
-\clearpage
 
@@
 \subsection{Descriptive Statistics}
+To get an overview of the data, the following descriptive statistics are presented.
+Statistics of interest are related to discussion papers, JEL Codes, and authors.
+In Figure \ref{fig:dp_counts}, it can be seen that the number of discussion papers published per month has increased over time from less than 20 to over 80 per month.
+Interesting to observe is the positive impact of the initial period of the COVID-19 pandemic on the number of discussion papers published.
+The average number of JEL Codes per discussion paper on the other hand is rather stable over time, and fluctuates only mildly.
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_counts_avg_jel_codes}
-    \caption{Discussion Papers \& Avg. JEL Codes per Paper Over Time}
-    \label{fig:monthly_counts_and_avg_jel_codes_per_paper}
+    \includegraphics{../bld/figures/fig_dp_counts}
+    \caption{Number of Discussion Papers \& JEL Codes per Discussion Paper}
+    \label{fig:dp_counts}
 \end{figure}
+
+Relevant to note is, that the structure of the authors publishing discussion papers has changed over time.
+The average number of authors per discussion paper per month has increased by about 0.5 authors per discussion paper.
+Furthermore, the average new number of new authors per discussion paper per month is stable around 0.5 authors per discussion paper, while the average number of authors that have published before is increasing by about 1.5 authors per discussion paper per month.
+This fact can be related to the establishment of the discussion series as a well-known platform for research in Labor Economics.
+These trends are shown in Figure \ref{fig:author_trends}.
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_author_trends}
+    \includegraphics{../bld/figures/fig_author_trends}
     \caption{Average Number of Authors Over Time}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_dp_counts}
-    \caption{Number of Authors \& JEL Codes per Discussion Paper}
-    \label{fig:dp_counts}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_top5overall_jel}
-    \caption{Top 5 JEL Codes}
-    \label{fig:top5overall_jel}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_top3yearly_jel}
-    \caption{Top 3 JEL Codes yearly}
-    \label{fig:top3yearly_jel}
-\end{figure}
-
 \section{Methodology}
+The methodologies presented here are used to achieve two main objectives.
+First, the behavior of subfields over time is to be analyzed.
+Second, the text data methods, such as topic modeling using Latent Dirichlet Allocation (LDA), are used to validate the JEL Codes as a tool for analyzing specialization trends.
 
@@
 To analyze the specialization trends in economics research, topic modeling is used.
-Specifically, the Latent Dirichlet Allocation (LDA) algorithm is used to identify topics in the abstracts of the discussion papers \cite{blei2003}.
+Specifically, the LDA algorithm is used to identify topics in the abstracts of the discussion papers \cite{blei2003}.
 To prepare the abstract data for the LDA model, the text is lowercased, tokenized, and stemmed.
@@
 
+In Figure \ref{fig:top5_increasing_decreasing_topics}, the top 5 topics increasing and decreasing in presence are shown.
+
+\begin{figure}
+    \centering
+    \includegraphics{../bld/figures/fig_topic_trends}
+    \caption{Top 5 Topics Increasing (Left) and Decreasing (Right) in Presence}
+    \label{fig:top5_increasing_decreasing_topics}
+\end{figure}
+
 \subsection{JEL Code Analysis}
+
+\begin{figure}
+    \centering
+    \includegraphics{../bld/figures/fig_top5overall_jel}
+    \caption{Top 5 JEL Codes}
+    \label{fig:top5overall_jel}
+\end{figure}
+
+In the Figure \ref{fig:top3yearl_jel}, all JEL Codes are shown that are at least in one year among the top 3 JEL Codes.
+If the JEL Code is not among the top 3 JEL Codes in a year, the opacity of the line is reduced.
+The JEL Codes 'J24' (Human Capital, Skills, Occupational Choice, Labor Productivity) and 'J31' (Wage Level and Structure, Wage Differentials) are consistently among the top 3 JEL Codes.
+While other JEL Codes like 'J16' (Economics of Gender, Non-labor Discrimination) saw a relative frequency increase lately.
+
+\begin{figure}
+    \centering
+    \includegraphics{../bld/figures/fig_top3yearly_jel}
+    \caption{Top 3 JEL Codes yearly}
+    \label{fig:top3yearly_jel}
+\end{figure}
 
@@
 
-\clearpage
-
 \section{Conclusion}
@@
 
+\section{Appendix}
+
+There are few authors that have published a large number of discussion papers.
+Similarly, there are few JEL Codes that are assigned to a large number of discussion papers.
+The Figure \ref{fig:counts_dp_jel_codes} shows the distribution of discussion papers by author on the left and by JEL Codes on the right.
+
+\begin{figure}
+    \centering
+    \includegraphics{../bld/figures/fig_counts_dp_jel_codes}
+    \caption{Discussion Papers by Author \& Discussion Papers by JEL Codes}
+    \label{fig:counts_dp_jel_codes}
+\end{figure}
+
 \end{document}

### 2025-02-27 documents/task.py

@@ -9,11 +9,12 @@ from pytask_latex import compilation_steps as cs
 from econ_spec_jel.config import BLD, DOCUMENTS, ROOT
 
 figures = [
-    "fig_counts_avg_jel_codes.png",
+    "fig_counts_dp_jel_codes.png",
     "fig_author_trends.png",
     "fig_dp_counts.png",
     "fig_top5overall_jel.png",
     "fig_top3yearly_jel.png",
+    "fig_topic_trends.png",
 ]
 DOCUMENTS_KWARGS = {
     "paper": {"depends_on": [BLD / "figures" / figure for figure in figures]},

### 2025-02-27 src/econ_spec_jel/analysis/task_network.py
Remove network analysis remainders.

@@ -1,55 +1,0 @@
-"""Task to create network of JEL codes."""
-
-import networkx as nx
-
-from typing import Annotated
-from pathlib import Path
-import pandas as pd
-
-from econ_spec_jel.config import DATACATALOGS
-
-
-def task_create_network(
-    data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
-) -> None:
-    """Create network of JEL codes."""
-    co_occurrence = _create_co_occurrence_df(data)
-    _create_graph_network(co_occurrence)
-
-
-def _create_co_occurrence_df(data: pd.DataFrame) -> pd.DataFrame:
-    exploded = data.explode("jel_codes")
-    merged = exploded.merge(exploded, on="dp_number")
-    co_occurrence_df = (
-        merged[merged["jel_codes_x"] != merged["jel_codes_y"]]
-        .groupby(["jel_codes_x", "jel_codes_y"])
-        .size()
-        .reset_index(name="co_occurrence")
-    )
-    return co_occurrence_df.pivot_table(
-        index="jel_codes_x", columns="jel_codes_y", values="co_occurrence", fill_value=0
-    )
-
-
-def _create_graph_network(co_occurrence_matrix: pd.DataFrame) -> nx.Graph:
-    graph = nx.Graph()
-    edges = _create_edges(co_occurrence_matrix)
-    graph.add_weighted_edges_from(edges)
-    return graph
-
-
-def _create_edges(co_occurrence_matrix: pd.DataFrame) -> list:
-    edges = co_occurrence_matrix.reset_index().melt(
-        id_vars="jel_codes_x", var_name="jel_codes_y", value_name="weight"
-    )
-    non_zero_and_non_looping_edges = edges[
-        (edges["weight"] > 0) & (edges["jel_codes_x"] != edges["jel_codes_y"])
-    ]
-    non_zero_and_non_looping_edges["weight"] = (
-        non_zero_and_non_looping_edges["weight"] * 10
-    )
-    return (
-        non_zero_and_non_looping_edges[["jel_codes_x", "jel_codes_y", "weight"]]
-        .to_records(index=False)
-        .tolist()
-    )

### 2025-02-27 documents/paper.tex
Adapt width of figure and remove comments from text_model task.

@@
     \centering
-    \includegraphics{../bld/figures/fig_dp_counts}
+    \includegraphics[width=\textwidth]{../bld/figures/fig_dp_counts}
     \caption{Number of Discussion Papers \& JEL Codes per Discussion Paper}
@@
     \centering
-    \includegraphics{../bld/figures/fig_author_trends}
+    \includegraphics[width=\textwidth]{../bld/figures/fig_author_trends}
     \caption{Average Number of Authors Over Time}
@@
     \centering
-    \includegraphics{../bld/figures/fig_topic_trends}
+    \includegraphics[width=\textwidth]{../bld/figures/fig_topic_trends}
     \caption{Top 5 Topics Increasing (Left) and Decreasing (Right) in Presence}
@@
     \centering
-    \includegraphics{../bld/figures/fig_top5overall_jel}
+    \includegraphics[width=\textwidth]{../bld/figures/fig_top5overall_jel}
     \caption{Top 5 JEL Codes}
@@
     \centering
-    \includegraphics{../bld/figures/fig_top3yearly_jel}
+    \includegraphics[width=\textwidth]{../bld/figures/fig_top3yearly_jel}
     \caption{Top 3 JEL Codes yearly}
@@
     \centering
-    \includegraphics{../bld/figures/fig_counts_dp_jel_codes}
+    \includegraphics[width=\textwidth]{../bld/figures/fig_counts_dp_jel_codes}
     \caption{Discussion Papers by Author \& Discussion Papers by JEL Codes}

### 2025-02-27 src/econ_spec_jel/text_model/task_model.py

@@ -8,7 +8,6 @@ from gensim.corpora import Dictionary
 from gensim.models import LdaModel
 
 
-# @pytask.mark.skip()
 def task_gensim_dictionary(
     data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
 ) -> Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]]:
@@ -25,7 +24,6 @@ def task_gensim_dictionary(
     return _get_gensim_dictionary(data["abstract_tokenized"])
 
 
-# @pytask.mark.skip()
 def task_corpus(
     data: Annotated[Path, DATACATALOGS["data"]["analysis"]],
     gensim_dictionary: Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]],
@@ -44,7 +42,6 @@ def task_corpus(
     return [gensim_dictionary.doc2bow(doc) for doc in data["abstract_tokenized"]]
 
 
-# @pytask.mark.skip()
 def task_train_topic_model(
     gensim_dictionary: Annotated[Path, DATACATALOGS["topic_model"]["dictionary"]],
     corpus: Annotated[Path, DATACATALOGS["topic_model"]["corpus"]],

### 2025-02-28 documents/paper.tex
Literarture review and conclusion.

@@
 % submissions.
-\usepackage[backend=bibtex, style=authoryear]{biblatex}
+\usepackage[backend=bibtex, style=authoryear, natbib]{biblatex}
 
@@
 \maketitle
-\footnotetext[1]{Email: \href{mailto:s87fschm@uni-bonn.de}{s87fschm@uni-bonn.de}, Matriculation Number: 50173514}
+\footnotetext[1]{Email: \href{mailto:s87fschm@uni-bonn.de}{s87fschm@uni-bonn.de}, Matriculation Number: 50173514, For replication material see \href{https://github.com/felixschmitz/econ_spec_jel}{here}}
 
 \section{Introduction}
-This paper proposes research to investigates specialization trends within economics research, focusing on whether subfields are becoming more specialized over time.
+This paper proposes research to investigate specialization trends within economics research, focusing on whether subfields are becoming more specialized over time.
 Understanding these trends is crucial as increasing specialization can influence knowledge dissemination, collaboration patterns, and the evolution of economic thought.
@@
 The JEL Codes are assigned by authors or editors and are intended to reflect the main topics of a paper.
-While they do not cover the methods or data used in a paper, they provide a useful starting point for understanding the content of a paper, and hence of economic research in general.
+While they do not cover the methods or data used in a paper, they provide a useful starting point for understanding the content of a paper, and hence of economic research in aggregate.
 By examining the intersections of JEL Codes across papers, this study assesses the depth and diversity of specialization within economics research.
@@
 This research also aims to validate the relevance of JEL Codes as a tool for analyzing specialization trends.
-Additionally, as an empirical application, the goal is to explores specialization dynamics within the field of Labor Economics, providing a focused example of how specialization may evolve over time.
-Through this approach, the study contributes to a better understanding of how economic research is structured and how subfields are connected or fragmented.
+Additionally, as an empirical application, the goal is to explore specialization dynamics within the subfield of Labor Economics, providing a focused example of how specialization may evolve over time.
+Through this approach, the study contributes to a better understanding of how economic research is structured and how subfields are connected or fragmented (e.g. \citep{davis2019}).
 The findings may have implications for how economists engage in interdisciplinary work and how academic communities organize knowledge.
@@
 The following sections of this research proposal are structured as follows.
-First, the literature review provides an overview of the existing research on specialization trends in economics and the use of JEL Codes.
+First, the literature review provides a brief overview of the existing research on specialization trends in economics and the use of JEL Codes.
 Secondly, the data used for the analysis is described, including the data collection process and pre-processing steps.
@@
 The literature on specialization trends in economics research is limited, with few studies focusing on the evolution of subfields within the discipline.
-However, there is a growing interest in understanding how knowledge is structured and how research communities are organized \cite{heikkila2022a}.
-investigates the use of JEL codes in economics research, and their development over time.
+\citet{onder2021} analyze the specialization trends of authors and co-authors in economics research, and the resulting impact on research output.
+They find that eventhough specialization of authors is increasing, the publication by them into highly ranked journals is decreasing.
+Recent advances in text data analysis have opened up new possibilities for studying specialization trends \citep{anauati2016, angrist2017, angrist2020, galiani2023a}.
+These facilitate natural language processing techniques to assign specific field of economics research tags.
+The analysis then is based upon these tags to identify specialization trends on the distinction of theoretical vs. empirical developments.
+The tags not only contain information on the topics covered, but also the approach and methods used in the research.
+\citet{heikkila2022a} investigates the use of JEL codes in economics research, and their development over time.
 The study finds that JEL codes are widely used to classify research in economics, and highlights the potential for "bibliometric and scientometric research" using these codes.
+In addition, the research shows that classification scheme of JEL Codes remained basically unchanged since a large revision in 1991.
+There is further research on the classification of research papers in economics between human and machine classification \citep{heikkila2024}.
+Generating (artificial) JEL Codes from abstracts using large language models is the focus here.
+There is relevant research on the trends of JEL Codes used in economics research \citep{kelly2011}.
+The authors analyze the publication by subfields, the development of JEL Codes, and define specialty journals using JEL codes.
+\citet{rath2016} extend this research focusing on the distribution of JEL Codes in economics research per subfield.
+Further, there is a growing interest in understanding how knowledge is structured and how research communities are organized.
+
+This research paper aims to contribute to the existing literature by validating the use of JEL Codes as a tool for analyzing specialization trends in economics research.
+In addition, the study will add a different perspective through the use of JEL Codes instead of predicted fields of economics research tags, and allow for a perspective on how authors themselves classify their research.
+To be more precise, the self-classification of JEL Codes and the subsequent dynamics might differ from the classification based upon text data based methods.
 
@@
 The data was scraped from the IZA website using the Python library \texttt{requests} and modules from the 'html-parser-library' \texttt{bs4.BeautifulSoup}.
-For each discussion paper, IZA has a webpage with metadata and a link to the PDF file.
-Both the metadata and the PDF file were downloaded and stored locally.
-The metadata for each discussion paper contains a title, a list of authors, the publication date (year and month), the list of assigned JEL Codes, a list of self-assigned keywords, and an abstract, among others.
+For each discussion paper, IZA has a webpage with metadata and a link to the discussion paper file.
+Both the metadata and the file were downloaded and stored locally.
+The metadata for each discussion paper contains a title, a list of authors, the publication date (year and month), the list of assigned JEL Codes, a list of assigned keywords, and an abstract, among others.
 It was also possible to extract links provided for individual authors, which, if they exist, reference an IZA profile page with further information on the author, or a personal website.
-The PDF files were not used for the analysis so far.
+The files and keywords were not used for the analysis so far.
 For this analysis, data has been collected up to February 2025 (discussion paper number 17695).
@@
 The later five ones fall into the time span January 2019 to November 2023.
-It is not possible to tell whether these discussion papers have been removed from the IZA website or whether they have never been published.
+It is not possible to tell why and when these discussion papers have been removed by IZA.
 
@@
     \label{fig:dp_counts}
-\end{figure}
-
-Relevant to note is, that the structure of the authors publishing discussion papers has changed over time.
-The average number of authors per discussion paper per month has increased by about 0.5 authors per discussion paper.
-Furthermore, the average new number of new authors per discussion paper per month is stable around 0.5 authors per discussion paper, while the average number of authors that have published before is increasing by about 1.5 authors per discussion paper per month.
-This fact can be related to the establishment of the discussion series as a well-known platform for research in Labor Economics.
-These trends are shown in Figure \ref{fig:author_trends}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../bld/figures/fig_author_trends}
-    \caption{Average Number of Authors Over Time}
-    \label{fig:author_trends}
 \end{figure}
@@
 
-In the Figure \ref{fig:top3yearl_jel}, all JEL Codes are shown that are at least in one year among the top 3 JEL Codes.
-If the JEL Code is not among the top 3 JEL Codes in a year, the opacity of the line is reduced.
-The JEL Codes 'J24' (Human Capital, Skills, Occupational Choice, Labor Productivity) and 'J31' (Wage Level and Structure, Wage Differentials) are consistently among the top 3 JEL Codes.
-While other JEL Codes like 'J16' (Economics of Gender, Non-labor Discrimination) saw a relative frequency increase lately.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../bld/figures/fig_top3yearly_jel}
-    \caption{Top 3 JEL Codes yearly}
-    \label{fig:top3yearly_jel}
-\end{figure}
-
 \subsection{Overlapping Network Analysis}
@@
 In a next step, the self-assigned keywords of the discussion papers can be used for comparison with the results from the topic modeling and JEL Code analysis.
+Further, the content of the files can be used to create artificial JEL Codes as in \citet{heikkila2024} or to create fields of economics research tags as shown by \citet{anauati2016, galiani2023a}.
+Leveraging the data from \citet{anauati2016, galiani2023a} would also allow to expand the analysis into other subfields of economics research, and to combine the JEL Code analysis with citation data.
+This would allow to understand how specialization trends in economics research are related to citation patterns and the dissemination of knowledge for specific topics.
 
 \printbibliography
+
+\clearpage
 
@@
 
+Relevant to note is, that the structure of the authors publishing discussion papers has changed over time.
+The average number of authors per discussion paper per month has increased by about 0.5 authors per discussion paper.
+Furthermore, the average new number of new authors per discussion paper per month is stable around 0.5 authors per discussion paper, while the average number of authors that have published before is increasing by about 1.5 authors per discussion paper per month.
+This fact can be related to the establishment of the discussion series as a well-known platform for research in Labor Economics.
+These trends are shown in Figure \ref{fig:author_trends}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../bld/figures/fig_author_trends}
+    \caption{Average Number of Authors Over Time}
+    \label{fig:author_trends}
+\end{figure}
+
+In Figure \ref{fig:top3yearl_jel}, all JEL Codes are shown that are at least in one year among the top 3 JEL Codes.
+If the JEL Code is not among the top 3 JEL Codes for a specific year in the plot, the opacity of the line is reduced.
+The JEL Codes 'J24' (Human Capital, Skills, Occupational Choice, Labor Productivity) and 'J31' (Wage Level and Structure, Wage Differentials) are consistently among the top 3 JEL Codes.
+While other JEL Codes like 'J16' (Economics of Gender, Non-labor Discrimination) saw a relative frequency increase lately.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../bld/figures/fig_top3yearly_jel}
+    \caption{Top 3 JEL Codes yearly}
+    \label{fig:top3yearly_jel}
+\end{figure}
+
 \end{document}

### 2025-02-28 src/econ_spec_jel/analysis/plotting_helper.py
Fix plots.

@@ -276,6 +276,16 @@ def plot_most_common_jel_codes(
         number_of_codes (int): The number of most common JEL codes to plot.
     """
     most_common_codes = _get_most_common_codes(data=data, number=number_of_codes)
+
+    # Choose a color palette from Plotly
+    color_palette = px.colors.qualitative.Plotly
+    # Extend the palette if there are more codes than colors in the palette
+    extended_palette = (
+        color_palette * ((len(most_common_codes) // len(color_palette)) + 1)
+    )[: len(most_common_codes)]
+    # Map each JEL code to a unique color
+    color_mapping = dict(zip(most_common_codes, extended_palette, strict=False))
+
     fig = go.Figure()
 
     total_counts = (
@@ -295,25 +305,32 @@ def plot_most_common_jel_codes(
         )
 
         smooth_indices = jel_counts.index[::6]
+
+        # Use consistent color for all traces related to this JEL code
+        color = color_mapping[code]
+
+        # Monthly trace
         fig.add_trace(
             go.Scatter(
                 x=jel_counts["publication_year_month"],
                 y=jel_counts[f"{code}_normalized"],
                 mode="lines",
                 name=f"{code} (monthly)",
-                line={"width": 1, "dash": "dot"},
+                line={"color": color, "width": 1, "dash": "dot"},
                 opacity=0.3,
             )
         )
+        # 6-month average trace
         fig.add_trace(
             go.Scatter(
                 x=jel_counts.loc[smooth_indices, "publication_year_month"],
                 y=jel_counts.loc[smooth_indices, f"{code}_normalized_smooth"],
                 mode="lines+markers",
                 name=f"{code} (6-month avg.)",
-                line={"width": 2},
+                line={"color": color, "width": 2},
             )
         )
+
     fig.update_layout(
         template="plotly_white",
         xaxis={

### 2025-02-28 src/econ_spec_jel/analysis/task_topics_plot.py

@@ -50,7 +50,7 @@ def _plot_topic_trends(
         ],
         shared_xaxes="columns",
         shared_yaxes="all",
-        vertical_spacing=0.01,
+        vertical_spacing=0.05,
         horizontal_spacing=0.05,
     )
 
@@ -99,8 +99,8 @@ def _plot_topic_trends(
 
     fig.update_layout(
         template="plotly_white",
-        width=1500,
-        height=1200,
+        width=1200,
+        height=900,
         showlegend=False,
     )
     fig.update_xaxes(tickformat="%Y")

### 2025-02-28 documents/paper.tex
Finalize research proposal.

@@
 To analyze specialization, a robust measure is required.
-The focus will be on leveraging Journal of Economic Literature (JEL) Codes, which are widely used by economists to classify research.
-These codes provide a structured way to categorize papers into subfields, allowing for a systematic analysis of specialization trends.
-The JEL Codes are assigned by authors or editors and are intended to reflect the main topics of a paper.
+Leveraging Journal of Economic Literature (JEL) Codes, which are widely used by economists to classify research, will be the aim of this research.
+JEL Codes provide a structured way to categorize papers into subfields, allowing for a systematic analysis of specialization trends.
+These are assigned by authors or editors and are intended to reflect the main topics of a paper.
 While they do not cover the methods or data used in a paper, they provide a useful starting point for understanding the content of a paper, and hence of economic research in aggregate.
@@
 Additionally, as an empirical application, the goal is to explore specialization dynamics within the subfield of Labor Economics, providing a focused example of how specialization may evolve over time.
-Through this approach, the study contributes to a better understanding of how economic research is structured and how subfields are connected or fragmented (e.g. \citep{davis2019}).
-The findings may have implications for how economists engage in interdisciplinary work and how academic communities organize knowledge.
+Through this approach, the study contributes to a better understanding of how economic research is structured and how subfields are connected or fragmented (see e.g. \citet{davis2019}).
+Findings may have implications for how economists engage in interdisciplinary work and how academic communities organize knowledge.
 
@@
 \citet{onder2021} analyze the specialization trends of authors and co-authors in economics research, and the resulting impact on research output.
-They find that eventhough specialization of authors is increasing, the publication by them into highly ranked journals is decreasing.
+They find that even though specialization of authors is increasing, the publication by them into highly ranked journals is decreasing.
 Recent advances in text data analysis have opened up new possibilities for studying specialization trends \citep{anauati2016, angrist2017, angrist2020, galiani2023a}.
@@
 The tags not only contain information on the topics covered, but also the approach and methods used in the research.
-\citet{heikkila2022a} investigates the use of JEL codes in economics research, and their development over time.
-The study finds that JEL codes are widely used to classify research in economics, and highlights the potential for "bibliometric and scientometric research" using these codes.
+\citet{heikkila2022a} investigates the use of JEL Codes in economics research, and their development over time.
+The study finds that JEL Codes are widely used to classify research in economics, and highlights the potential for "bibliometric and scientometric research" using these Codes.
 In addition, the research shows that classification scheme of JEL Codes remained basically unchanged since a large revision in 1991.
@@
 There is relevant research on the trends of JEL Codes used in economics research \citep{kelly2011}.
-The authors analyze the publication by subfields, the development of JEL Codes, and define specialty journals using JEL codes.
+The authors analyze the publication by subfields, the development of JEL Codes, and define specialty journals using JEL Codes.
 \citet{rath2016} extend this research focusing on the distribution of JEL Codes in economics research per subfield.
@@
 For a future project, it would be interesting to extend the analysis to other working paper series or general economics journals.
-A dataset that might be easily available for analysis is a combination of Constellate, Semantic Scholar, and manually classified publications into subfields of economics research, as produced by \cite{galiani2023a}.
+A dataset that might be easily available for analysis is a combination of Constellate, Semantic Scholar, and manually classified publications into subfields of economics research, as produced by \citet{galiani2023a}.
 This dataset would provide a more comprehensive view of specialization trends in economics research, and would allow to join the work with .
@@
 The letters were converted to uppercase, leading '1' and '0' were replaced by 'I' and 'O', respectively.
-This confusion is likely due to the extraction of the JEL Codes via Optical Character Recognition (OCR) from the PDFs.
-After these thoroughly cleaning steps, JEL Codes not starting with a capital letter followed followed by at least one and at most two digits were dropped.
+This confusion is likely due to the extraction of the JEL Codes via Optical Character Recognition from the files.
+After these thoroughly cleaning steps, JEL Codes not starting with a capital letter followed by at least one and at most two digits were dropped.
 Finally, JEL Codes with a single digit were converted to two digits by adding a trailing zero.
@@
 This is mainly to ensure a sufficient number of observations per month (for the early years) and complete information on the discussion paper (for the most recent discussion papers).
-This reduces the sample size by 195 observations. The final number of discussion papers analysed is 17,449.
+This reduces the sample size by 195 observations. The final number of discussion papers analyzed is 17,449.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../bld/figures/fig_dp_counts}
+    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_dp_counts}
     \caption{Number of Discussion Papers \& JEL Codes per Discussion Paper}
@@
 
+In Figure \ref{fig:top5overall_jel}, the overall top 5 JEL Codes are shown as the normalized number of appearance per month and over the smoothed 6-month-period.
+The normalization is done by dividing the number of discussion papers with a JEL Code by the total number of discussion papers in the respective month.
+The smoothing is done by taking the average of a rolling window of 6 months.
+A spike in the normalized frequency of the JEL Code 'J31' (Wage Level and Structure, Wage Differentials) is likely related to an employment-related collective bargaining in Germany \citep{institutfurarbeitsrechtundarbeitsbeziehungenindereuropaischenunioniaaeuuniversitattrier2001}, when the discussion paper series was still young, and likely used frequently by German authors.
+After this event, the frequency of the JEL Code 'J31' decreased again, and almost vanishes over time.
+This is contrasted to JEL Codes 'J16' (Economics of Gender, Non-labor Discrimination) and 'J24' (Human Capital, Skills, Occupational Choice, Labor Productivity), which gained substantially in frequency over time.
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_top5overall_jel}
+    \caption{Top 5 JEL Codes}
+    \label{fig:top5overall_jel}
+\end{figure}
+
 \section{Methodology}
-The methodologies presented here are used to achieve two main objectives.
-First, the behavior of subfields over time is to be analyzed.
-Second, the text data methods, such as topic modeling using Latent Dirichlet Allocation (LDA), are used to validate the JEL Codes as a tool for analyzing specialization trends.
+The methodology for analyzing specialization trends in economics research is based on two main approaches: topic modeling and network analysis.
+These methods are used to identify thematic structures within the corpus of discussion papers and JEL Codes assigned to analyze specialization trends.
 
 \subsection{Topic Modeling}
-To analyze the specialization trends in economics research, topic modeling is used.
-Specifically, the LDA algorithm is used to identify topics in the abstracts of the discussion papers \cite{blei2003}.
-To prepare the abstract data for the LDA model, the text is lowercased, tokenized, and stemmed.
-Numbers, punctuation, and words containing no alphanumeric characters are removed.
-Furthermore, stopwords are removed from the text.
-The model
-The number of topics is set to 750 and 20 passes through the corpus of tokenized and stemmed abstracts are performed during training of the model, similar to \cite{galiani2023a}.
-
-\subsection{Topic Modeling}
-
 To analyze specialization trends in economics research, topic modeling is employed to identify latent topics within the abstracts of discussion papers.
-Specifically, the Latent Dirichlet Allocation (LDA) generative model is used, which assumes that documents discussing similar topics tend to use similar groups of words \cite{blei2003}.
+Specifically, the Latent Dirichlet Allocation (LDA) generative model is used, which assumes that documents discussing similar topics tend to use similar groups of words \citep{blei2003}.
 LDA is particularly suitable for this analysis as it captures co-occurring word clusters, revealing thematic structures within the corpus.
@@
 Stopwords, which are common function words with little semantic value, are also eliminated.
+A corpus of words is then created, and words that appear in less than 20 documents and words that appear in more than 90\% of documents are removed.
+This corpus is then transformed into a document-term matrix, where each row represents a document and each column a word.
 These preprocessing steps collectively improve the quality of the tokenized text, ensuring a more effective training process for the LDA model.
 
-The LDA model is trained using \texttt{Gensim}'s implementation of the LDA training algorithm
-%\cite{rehurek2010}.
+The LDA model is trained using \texttt{Gensim}'s implementation of the LDA training algorithm \citep{rehurek2010}.
 The number of topics is set to 750, and the model is trained using 20 passes through the corpus.
-These settings are chosen to ensure sufficient topic granularity while maintaining interpretability, similar to the approach used in \cite{galiani2023a}.
-
-% is this true?
-The model uses a symmetric Dirichlet prior for the topic distribution per document ($\alpha$) and the word distribution per topic ($\eta$), reflecting the assumption that all topics and words are equally likely a priori.
+These settings are chosen to ensure sufficient topic granularity while maintaining interpretability, similar to the model specifications used in \citet{galiani2023a}.
 
@@
 
-To ensure the meaningfulness of the detected topics, a filtering process is applied.
-Topics predominantly composed of non-thematic words (e.g., "online", "access", "appendix") or those lacking a clear thematic pattern are excluded.
-% After this filtering process, 695 meaningful topics are retained.
-An article $i$ is defined as "containing" topic $k$ if $t_{i,k}$ ranks among the top ten topics with the highest presence in that article.
-This approach ensures consistent topic assignment while maintaining interpretability.
+In Figure \ref{fig:top5_increasing_decreasing_topics}, the top 5 topics increasing and decreasing in presence are shown.
+The left column contains topics with the highest increase in presence, while the right column contains topics with the lowest increase in presence (highest decrease in presence).
+Plots are sorted in each column from by the value of the increase in presence, hence the topic with the highest increase in presence is shown in the left column and first row, and the most decreasing topic in right column and fifth row.
 
-This topic modeling approach not only identifies prevalent themes but also captures the interconnectedness of topics within the field.
-It provides a robust foundation for analyzing specialization trends in economics research by revealing how topics cluster and evolve over time.
-The results from this robust methodology offer a comprehensive benchmark for understanding the dynamics of specialization within the field when using JEL codes.
-
-In Figure \ref{fig:top5_increasing_decreasing_topics}, the top 5 topics increasing and decreasing in presence are shown.
-
-\begin{figure}
+\begin{figure}[H]
     \centering
@@
 
-\subsection{JEL Code Analysis}
+The most increasing topic contains stemmed words such as ("student", "academ"), making it related to research in education.
+The third most increasing topic contains stemmed words such as ("gender", "gap"), indicating research focusing on discrepancies between genders.
+The most decreasing topic contains stemmed words such as ("wage", "increas", higher"), indicating research on wage increases.
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../bld/figures/fig_top5overall_jel}
-    \caption{Top 5 JEL Codes}
-    \label{fig:top5overall_jel}
-\end{figure}
+This topic modeling approach not only identifies prevalent themes but also captures the interconnectedness of topics within the field.
+It provides a robust foundation for analyzing specialization trends in economics research by revealing how topics cluster and evolve over time.
+The results from this robust methodology offer a comprehensive benchmark for understanding the dynamics of specialization within the field and will be linked with the results from network analysis based on JEL Codes.
 
 \subsection{Overlapping Network Analysis}
-Initially, I wanted to use the Louvain algorithm to detect communities in the network.
-However, the algorithm only allows to non-overlapping communities, that is, a node can only belong to one community.
+Initially, the method selected to use was the Louvain algorithm to detect communities in the network.
+There, a network of nodes (unique JEL Codes) and edges (co-occurrence of JEL Codes in discussion papers) is created, and the algorithm is applied to dissect the network into non-overlapping communities.
+That is, a node can only belong to one community.
 This is not suitable for the analysis of specialization trends in economics research, as a JEL code can belong to multiple subfields.
-Rather the combination of JEL codes describes a subfield.
-Subsets of JEL codes describing a subfield might overlap with JEL codes describing another subfields.
-Hence, the goal is to use a different algorithm to detect overlapping communities in the network.
+Figure \ref{fig:top5overall_jel} supports this argument, as the JEL Code 'J61' (Geographic Labor Mobility, Immigrant Workers) is not perceived to undergo a relevant increase in presence, yet the topic is of high interest in the field of Labor Economics \citep{fasani2020}.
+Unique JEL Codes might be shared between subfields, and hence the communities are overlapping.
+This combination of JEL Codes then describes a subfield.
+Hence, the goal is to use an algorithm able to detect overlapping communities in the network, such as described in \citet{yang2013}.
 
 \section{Conclusion}
-In a next step, the self-assigned keywords of the discussion papers can be used for comparison with the results from the topic modeling and JEL Code analysis.
-Further, the content of the files can be used to create artificial JEL Codes as in \citet{heikkila2024} or to create fields of economics research tags as shown by \citet{anauati2016, galiani2023a}.
-Leveraging the data from \citet{anauati2016, galiani2023a} would also allow to expand the analysis into other subfields of economics research, and to combine the JEL Code analysis with citation data.
-This would allow to understand how specialization trends in economics research are related to citation patterns and the dissemination of knowledge for specific topics.
+This research examines specialization trends within economics using JEL Codes, employing topic modeling and network analysis to identify thematic structures and specialization dynamics.
+The findings provide insights into how economic research is structured and how subfields are interconnected or fragmented.
+Next steps involve comparing self-assigned keywords with results from topic modeling and JEL Code analysis.
+Additionally, the content of the files can be used to generate artificial JEL Codes \citep{heikkila2024} or to create research tags following methods by \citet{anauati2016, galiani2023a}.
+Expanding the analysis with data from \citet{anauati2016, galiani2023a} would enable exploration of other subfields and integration with citation data.
+This would enhance understanding of how specialization trends relate to citation patterns and knowledge dissemination within economic research.
+Through these extensions, the study aims to provide valuable perspectives on the evolution of economic specialization and its implications for interdisciplinary collaboration and knowledge organization.
+\clearpage
 
@@
 
-\begin{figure}
+\begin{figure}[H]
     \centering
-    \includegraphics[width=\textwidth]{../bld/figures/fig_counts_dp_jel_codes}
+    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_counts_dp_jel_codes}
     \caption{Discussion Papers by Author \& Discussion Papers by JEL Codes}
@@
 
-Relevant to note is, that the structure of the authors publishing discussion papers has changed over time.
+It is relevant to note that the structure of the authors publishing discussion papers has changed over time.
 The average number of authors per discussion paper per month has increased by about 0.5 authors per discussion paper.
@@
 
-\begin{figure}
+\begin{figure}[H]
     \centering
-    \includegraphics[width=\textwidth]{../bld/figures/fig_author_trends}
+    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_author_trends}
     \caption{Average Number of Authors Over Time}
@@
 
-In Figure \ref{fig:top3yearl_jel}, all JEL Codes are shown that are at least in one year among the top 3 JEL Codes.
+In Figure \ref{fig:top3yearly_jel}, all JEL Codes are shown that are at least in one year among the top 3 JEL Codes.
 If the JEL Code is not among the top 3 JEL Codes for a specific year in the plot, the opacity of the line is reduced.
@@
 
-\begin{figure}
+\begin{figure}[H]
     \centering
-    \includegraphics[width=\textwidth]{../bld/figures/fig_top3yearly_jel}
+    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_top3yearly_jel}
     \caption{Top 3 JEL Codes yearly}
